### 2019-03-26 full analysis file.Rmd
Fixed paths and other stuff

---
title: "Objectification Special Studies Analysis"
author: "Kat Kyuchukova"
date: "9/18/2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(nlme)
library(haven)
library(devtools)
#install_github("RandiLGarcia/dyadr")
#library(dyadr)
library(psych)
library(mosaic)
library(tidyr)
library(dplyr)
library(irr)
library(lavaan)
select <- dplyr::select
```

Load data
```{r}
#clark
clark_measures <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Measures_1.sav")
clark_pairwise <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Pairwise_1.sav")

#smith
smith <- read_spss("Raw data/Smith College Relationships Data/College_Relationship_Dyadic_September_14_2018_12_17.sav")

#sentence data

clark_sentences <- read.csv("Raw data/Clark College Relationships Data/sexist sentences/coded_sentences.csv")

#trigrams
trigram_check <- read.csv("Raw data/Clark College Relationships Data/trigrams/AMELIA_check_trigram_grading.csv")
```

#looking at variables
```{r}
glimpse(clark_measures)
glimpse(clark_pairwise)
names(clark_measures)
```

CLARK SCALE SCORES
```{r}
#Social Physique Anxiety
clark_measures <- clark_measures %>%
  mutate(physanx5.r = (min(physanx_5, na.rm = TRUE) + max(physanx_5, na.rm = TRUE) - physanx_5),
         physanx = (physanx_1 +physanx_2 +physanx_3+physanx_4+physanx5.r+physanx_6)/6)

#SOO
clark_measures$extSOO <- rowMeans(dplyr::select(clark_measures, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE)

clark_measures$intSOO <- rowMeans(dplyr::select(clark_measures, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE)

clark_measures$a2pSOO <- clark_measures$extSOO - clark_measures$intSOO

psych::alpha(select(clark_measures, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE)

psych::alpha(select(clark_measures, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE)

#Meta SOO
clark_measures$extmetaSOO <- rowMeans(dplyr::select(clark_measures, metaint_1, metaint_2, metaint_5, metaint_6, metaint_10), na.rm=TRUE)

clark_measures$intmetaSOO <- rowMeans(dplyr::select(clark_measures, metaint_3, metaint_4, metaint_7, metaint_8, metaint_9, metaint_11, metaint_12), na.rm=TRUE)

clark_measures$metaSOO <- clark_measures$extmetaSOO - clark_measures$intmetaSOO

#Interaction Authenticity
clark_measures$authenticity <- rowMeans(dplyr::select(clark_measures, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

psych::alpha(select(clark_measures, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

#SSO
clark_measures <- clark_measures %>%
  mutate(howbody_avg = (howbody_1+howbody_2)/2)

#Relationship Agency
clark_measures$relag <- rowMeans(dplyr::select(clark_measures, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

psych::alpha(select(clark_measures, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

#Career Aspirations
clark_measures <- clark_measures %>%
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

clark_measures$careerasp <- rowMeans(dplyr::select(clark_measures, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

psych::alpha(select(clark_measures, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

#TSO
clark_measures$funcTSO <- rowMeans(dplyr::select(clark_measures,tso_1, tso_2, tso_4, tso_7, tso_9))
clark_measures$physTSO <- rowMeans(dplyr::select(clark_measures, tso_3, tso_5, tso_6, tso_8, tso_10))

clark_measures$TSO <- clark_measures$physTSO - clark_measures$funcTSO
clark_measures$TSO <- clark_measures$TSO*(-1)

#satisfaction--reverse code negative variables and then average with the rest
clark_measures <- clark_measures %>%
  mutate(anxious.r = (min(inter_2, na.rm = TRUE) + max(inter_2, na.rm = TRUE) - inter_2)) %>%
  mutate(uncomfort.r = (min(inter_4, na.rm = TRUE) + max(inter_4, na.rm = TRUE) - inter_4)) %>%
  mutate(hostile.r = (min(inter_8, na.rm = TRUE) + max(inter_8, na.rm = TRUE) - inter_8)) %>%
  mutate(apprehen.r = (min(inter_11, na.rm = TRUE) + max(inter_11, na.rm = TRUE) - inter_11)) %>%
  mutate(aggressive.r = (min(inter_13, na.rm = TRUE) + max(inter_13, na.rm = TRUE) - inter_13))

clark_measures$satisfaction <- rowMeans(select(clark_measures, anxious.r, uncomfort.r, hostile.r, apprehen.r, aggressive.r, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_6), na.rm=TRUE)
```

SMITH SCALE SCORES
```{r}
smith <- smith %>%
  arrange(dyadID) %>%
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999)

#Social Physique Anxiety
smith <- smith %>%
  mutate(physanx5.r = (min(physanx_5, na.rm = TRUE) + max(physanx_5, na.rm = TRUE) - physanx_5),
         physanx = (physanx_1 +physanx_2 +physanx_3+physanx_4+physanx5.r+physanx_6)/6)

#SOO
smith$extSOO <- rowMeans(dplyr::select(smith, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE)

psych::alpha(select(smith, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE)
smith$intSOO <- rowMeans(dplyr::select(smith, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE)

psych::alpha(select(smith, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE)
smith$a2pSOO <- smith$extSOO - smith$intSOO

#Meta SOO
smith$extmetaSOO <- rowMeans(dplyr::select(smith, metaint_1, metaint_2, metaint_5, metaint_6, metaint_10), na.rm=TRUE)

smith$intmetaSOO <- rowMeans(dplyr::select(smith, metaint_3, metaint_4, metaint_7, metaint_8, metaint_9, metaint_11, metaint_12), na.rm=TRUE)

smith$metaSOO <- smith$extmetaSOO - smith$intmetaSOO

#Interaction Authenticity
smith$authenticity <- rowMeans(dplyr::select(smith, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

psych::alpha(select(smith, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

#SSO
smith <- smith %>%
  mutate(howbody_avg = (howbody_1+howbody_2)/2)

#Relationship Agency
smith$relag <- rowMeans(dplyr::select(smith, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

psych::alpha(select(smith, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

#Career Aspirations
smith <- smith %>%
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

smith$careerasp <- rowMeans(dplyr::select(smith, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

psych::alpha(select(smith, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

#TSO
smith$funcTSO <- rowMeans(dplyr::select(smith,tso_1, tso_2, tso_4, tso_7, tso_9))
smith$physTSO <- rowMeans(dplyr::select(smith, tso_3, tso_5, tso_6, tso_8, tso_10))

smith$TSO <- smith$physTSO-smith$funcTSO
smith$TSO <- smith$TSO*-1

#satisfaction--reverse code negative variables and then average with the rest
smith <- smith %>%
  mutate(anxious.r = (min(inter_2, na.rm = TRUE) + max(inter_2, na.rm = TRUE) - inter_2)) %>%
  mutate(uncomfort.r = (min(inter_4, na.rm = TRUE) + max(inter_4, na.rm = TRUE) - inter_4)) %>%
  mutate(hostile.r = (min(inter_8, na.rm = TRUE) + max(inter_8, na.rm = TRUE) - inter_8)) %>%
  mutate(apprehen.r = (min(inter_11, na.rm = TRUE) + max(inter_11, na.rm = TRUE) - inter_11)) %>%
  mutate(aggressive.r = (min(inter_13, na.rm = TRUE) + max(inter_13, na.rm = TRUE) - inter_13))

smith$satisfaction <- rowMeans(select(smith, anxious.r, uncomfort.r, hostile.r, apprehen.r, aggressive.r, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_6), na.rm=TRUE)
```

TRIGRAMS
```{r}

###CLARK
trigram_check <- trigram_check[,1:8]

trigram_check <- trigram_check %>%
  mutate(trigram_correct = ifelse(accurate.=="yes", Correct...1.0., 1))

trigram_clark <- trigram_check %>%
  group_by(partID) %>%
  summarise(trigram_score = sum(trigram_correct)) %>%
  dplyr::select(partID, trigram_score)

#merging
clark <- merge(x=clark_measures, y=trigram_clark, by="partID", all.x=TRUE)

###SMITH
smith_trigrams <- smith %>%
  dplyr::select(partID, dyadID, rat1_1.0, rat2_1.0,rat3_1.0, rat4_1.0,rat5_1.0, rat6_1.0, rat7_1.0, rat8_1.0, rat9_1.0, rat10_1) %>%
  arrange(dyadID, partID)

smith_trigrams_long <- gather(smith_trigrams, trigram, response, rat1_1.0:rat10_1)

smith_trigrams_long <- smith_trigrams_long %>%
  mutate(trigram = ifelse(trigram=="rat1_1.0", "QUACK:POND:WADDLE",
                          ifelse(trigram=="rat2_1.0", "ATHLETES:WEB:RABBIT",
                                 ifelse(trigram=="rat3_1.0", "HONEY:SWARM:STING",
                                        ifelse(trigram=="rat4_1.0", "MATCHES:SMOKE:BUSH",
                                               ifelse(trigram=="rat5_1.0", "CHOCOLATE:FORTUNE:TIN",
                                                      ifelse(trigram=="rat6_1.0", "CURIOSITY:NAP:WHISKERS",
                                                             ifelse(trigram=="rat7_1.0", "STALK:TRAINER:KING",
                                                                    ifelse(trigram=="rat8_1.0", "SURPRISE:LINE:BIRTHDAY",
                                                                           ifelse(trigram=="rat9_1.0", "RED:GO:CAR", 
                                                                                  ifelse(trigram=="rat10_1", "MOUSE : SHARP : BLUE", NA)))))))))))

smith_trigrams_long <- smith_trigrams_long %>%
  mutate(correct_response=ifelse(trigram=="QUACK:POND:WADDLE", "Duck",
                          ifelse(trigram=="ATHLETES:WEB:RABBIT", "Foot",
                                 ifelse(trigram=="HONEY:SWARM:STING","Bee",
                                        ifelse(trigram=="MATCHES:SMOKE:BUSH", "Fire", 
                                               ifelse(trigram=="CHOCOLATE:FORTUNE:TIN", "Cookie",
                                                      ifelse(trigram=="CURIOSITY:NAP:WHISKERS","Cat",
                                                             ifelse(trigram=="STALK:TRAINER:KING","Lion",
                                                                    ifelse(trigram=="SURPRISE:LINE:BIRTHDAY", "Party",
                                                                           ifelse(trigram=="RED:GO:CAR", "Stop",
                                                                                  ifelse(trigram=="MOUSE : SHARP : BLUE", "Cheese", NA))))))))))) %>%
  arrange(partID, dyadID)

smith_trigrams_long <- smith_trigrams_long[,c(1,2,3,5,4)]

#write.csv(smith_trigrams_long, "/Users/katerinakyuchukova/Downloads/kat/Smith College Relationships Data/trigrams_to_grade.csv", row.names=FALSE)

smith_trigrams_graded <- read.csv("Raw data/Smith College Relationships Data/trigrams_graded.csv")

smith_triscore <- smith_trigrams_graded %>%
  group_by(partID, dyadID) %>%
  summarise(trigram_score=sum(final_correct))%>%
  dplyr::select(partID, dyadID, trigram_score)

smith <- merge(x=smith, y=smith_triscore, x.all=TRUE)
```
sexist sentences
```{r}
#ICC
clark_sentences <- clark_sentences[,1:8]

clark_sentences <- clark_sentences %>%
  filter(!is.na(dyadID))

clark_sentences <- na.omit(clark_sentences)
clark_icc <- clark_sentences %>%
  select(Hannah, Jennie, Amelia)

icc(clark_icc, model="twoway", unit="average")

#per sentence
clark_icc1 <- clark_sentences %>%
  filter(Sentence==1) %>%
  select(Hannah, Jennie, Amelia)

clark_icc2 <- clark_sentences %>%
  filter(Sentence==2) %>%
  select(Hannah, Jennie, Amelia)

clark_icc3 <- clark_sentences %>%
  filter(Sentence==3) %>%
  select(Hannah, Jennie, Amelia)

clark_icc4 <- clark_sentences %>%
  filter(Sentence==4) %>%
  select(Hannah, Jennie, Amelia)

clark_icc5 <- clark_sentences %>%
  filter(Sentence==5) %>%
  select(Hannah, Jennie, Amelia)

clark_icc6 <- clark_sentences %>%
  filter(Sentence==6) %>%
  select(Hannah, Jennie, Amelia)

clark_icc7 <- clark_sentences %>%
  filter(Sentence==7) %>%
  select(Hannah, Jennie, Amelia)

clark_icc8 <- clark_sentences %>%
  filter(Sentence==8) %>%
  select(Hannah, Jennie, Amelia)

icc(clark_icc1, model="twoway", unit="average")
icc(clark_icc2, model="twoway", unit="average")
icc(clark_icc3, model="twoway", unit="average")
icc(clark_icc4, model="twoway", unit="average")
icc(clark_icc5, model="twoway", unit="average")
icc(clark_icc6, model="twoway", unit="average")
icc(clark_icc7, model="twoway", unit="average")
icc(clark_icc8, model="twoway", unit="average")

clark_test <- clark_sentences %>%
  filter(Sentence!=4) %>%
  select(Hannah, Jennie, Amelia)

icc(clark_test, model="twoway", unit="average")

clark_sentences <- clark_sentences %>%
  filter(Sentence!=4)

#good enough to get an average??

clark_sentence_avg <- clark_sentences %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

clark_logel <- clark_sentences %>%
  filter(Sentence==3|Sentence==5|Sentence==6) %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

clark_sexism <- merge(x=clark, y=clark_sentence_avg, by="partID", all.x=TRUE)
clark_final <- merge (x=clark_sexism, y=clark_logel, by="partID", all.x=TRUE)

#Smith

#Creating sexist sentence file
smith$dyadID <- as.numeric(smith$dyadID)
smith$partID <- as.numeric(smith$partID)
smith_sentences_long <- smith %>%
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999) %>%
  select(partID, dyadID, sexismpre_1:sexismpre_8) %>%
  arrange(dyadID)

smith_sentences_long <- gather(smith_sentences_long, sentence, target, sexismpre_1:sexismpre_8)

smith_sentences_long <- smith_sentences_long %>%
  mutate(stem=ifelse(sentence=="sexismpre_1", "Linda swatted at the flies", 
                     ifelse(sentence=="sexismpre_2", "Helen crammed for the test",
                            ifelse(sentence=="sexismpre_3", "Shirley asked for help getting home",
                                   ifelse(sentence=="sexismpre_4", "Sam read the newspaper",
                                          ifelse(sentence=="sexismpre_5", "Jenny went home to cook dinner",
                                                 ifelse(sentence=="sexismpre_6", "Katherine baby-sat the neighbor's kids",
                                                        ifelse(sentence=="sexismpre_7", "Tom bottle fed the baby",
                                                               ifelse(sentence=="sexismpre_8", "Bert changed the oil", NA))))))))) %>%
    mutate(sentence=ifelse(sentence=="sexismpre_1", 1, 
                     ifelse(sentence=="sexismpre_2", 2,
                            ifelse(sentence=="sexismpre_3", 3,
                                   ifelse(sentence=="sexismpre_4", 4,
                                          ifelse(sentence=="sexismpre_5", 5,
                                                 ifelse(sentence=="sexismpre_6", 6,
                                                        ifelse(sentence=="sexismpre_7", 7,
                                                               ifelse(sentence=="sexismpre_8", 8, NA)))))))))

smith_sentences_long <- smith_sentences_long[,c(1,2,3,5,4)]

smith_sentences_long <- smith_sentences_long %>%
  filter(target!="") %>%
  arrange(dyadID, partID)

#write.csv(smith_sentences_long, "/Users/katerinakyuchukova/Downloads/kat/Smith College Relationships Data/sexist_sentences_for_coding.csv", row.names = FALSE)

#coding correlations
kat_codes <- read.csv("Raw data/Smith College Relationships Data/kat_sentences_coded.csv")

xinyu_codes <- read.csv("Raw data/Smith College Relationships Data/xinyu_sentences.csv")

coded_sentences <- merge(x=kat_codes, y=xinyu_codes, all.x = TRUE)

smith_icc <- coded_sentences %>%
  select(kat_score, Xinyu)

icc(smith_icc, model="twoway", unit="average")

##TAKE TWOOOO
kat_codes2 <- read.csv("Raw data/Smith College Relationships Data/kat_recode.csv")

xinyu_codes2 <- read.csv("Raw data/Smith College Relationships Data/xinyu_recode.csv")

coded_sentences2 <- merge(x=kat_codes2, y=xinyu_codes2, all.x = TRUE)

smith_icc2 <- coded_sentences2 %>%
  filter(sentence!=4)

icc(smith_icc2, model="twoway", unit="average")

#making average file
smith_sentence_avg <- coded_sentences2 %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

smith_logel <- coded_sentences2 %>%
  filter(sentence==3|sentence==5|sentence==6) %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

smith_sexism <- merge(x=smith, y=smith_sentence_avg, by="partID", all.x=TRUE)
smith_final <- merge (x=smith_sexism, y=smith_logel, by="partID", all.x=TRUE)

```

Making Pairwise
```{r}

#Clark

clark<- clark_final %>%
  filter(dyadID!=4, dyadID!=14, dyadID!=19) %>%
  arrange(dyadID) %>%
  mutate(person=rep(1:2, times=22)) %>%
  select(dyadID, person, partID, gender, age, race, sexor, physanx, a2pSOO, metaSOO, trigram_score, authenticity, howbody_avg, relag, careerasp, TSO, avg_sexism, sexism_logel, satisfaction) %>%
  arrange(person)

tempA <- clark %>% 
  mutate(personE = ifelse(person == 1, -1, 1), 
         partnum = 1,
         person = ifelse(person == 1, "A", "P")) %>%
  gather(variable, value, partID:partnum)%>%
  unite(var_person, variable, person)%>%
  spread(var_person, value)

tempB <- clark %>% 
  mutate(personE = ifelse(person == 1, -1, 1), 
         partnum = 2,
         person = ifelse(person == 1, "P", "A")) %>%
  gather(variable, value, partID:partnum) %>%
  unite(var_person, variable, person) %>%
  spread(var_person, value)

clark_pair <- bind_rows(tempA, tempB) %>%
  arrange(dyadID)

rm(tempA, tempB)

head(clark_pair)

#Smith

#making these numeric
smith_final$dyadID <- as.numeric(smith_final$dyadID)
smith_final$partID <- as.numeric(smith_final$partID)

#checking weird dyads to see if there is post interaction information
smith_checK <- smith %>%
  filter(dyadID==16|is.na(dyadID)|dyadID==798797|dyadID==999|dyadID==999999|dyadID==9999)%>%
  arrange(dyadID) %>%
  select(dyadID, partID, otherint_1:otherint_12)

smith <- smith_final %>%
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999, dyadID!=7) %>%
  arrange(dyadID) %>%
  mutate(person=rep(1:2, times=23),
         partID= ifelse(dyadID==31 & person==2, 83, partID),
         dyadID= ifelse(dyadID==2 & partID==4|partID==5, 1, dyadID)) %>%
  select(dyadID, person, partID, gender, age, race, sexor, physanx, a2pSOO, metaSOO, trigram_score, authenticity, howbody_avg, relag, careerasp, TSO, avg_sexism, sexism_logel, satisfaction)  %>%
  arrange(person)

tempA <- smith %>% 
  mutate(personE = ifelse(person == 1, -1, 1), 
         partnum = 1,
         person = ifelse(person == 1, "A", "P")) %>%
  gather(variable, value, partID:partnum)%>%
  unite(var_person, variable, person) %>%
  spread(var_person, value)

tempB <- smith %>% 
  mutate(personE = ifelse(person == 1, -1, 1), 
         partnum = 2,
         person = ifelse(person == 1, "P", "A")) %>%
  gather(variable, value, partID:partnum) %>%
  unite(var_person, variable, person) %>%
  spread(var_person, value)

smith_pair <- bind_rows(tempA, tempB)

rm(tempA, tempB)

head(smith_pair)
```

Next steps: correlations, matrices etc
```{r}

#making variables numeric
ix <- 1:39
smith_pair[,ix] <- lapply(smith_pair[,ix], as.numeric)
clark_pair[,ix] <-lapply(clark_pair[,ix], as.numeric)

#same gender
samegender_clark <- clark_pair %>%
  filter(gender_A==gender_P)

samegender_smith <- smith_pair %>%
  filter(gender_A==gender_P)

samegender_smith[,ix] <- lapply(samegender_smith[,ix], as.numeric)
samegender_clark[,ix] <-lapply(samegender_clark[,ix], as.numeric)

#corr.test(select(samegender_smith, physanx_A, a2pSOO_A, metaSOO_A, authenticity_A, howbody_avg_A, relag_A, careerasp_A, TSO_A, avg_sexism_A, sexism_logel_A))
#corr.test(select(samegender_clark, physanx, a2pSOO, metaSOO, authenticity, howbody_avg, relag, careerasp, TSO, avg_sexism, sexism_logel))

#corr.test(select(samegender_clark, a2pSOO_A, a2pSOO_P, metaSOO_A, metaSOO_P, howbody_avg_A))
#corr.test(select(diffgender_clark, a2pSOO_A, a2pSOO_P, metaSOO_A, metaSOO_P, howbody_avg_A))
#corr.test(select(samegender_clark, authenticity_A, careerasp_A, relag_A))

diffgender_clark <- clark_pair %>%
  filter(gender_A!=gender_P)

#smith
corr.test(select(smith_final, physanx, a2pSOO, metaSOO, authenticity, howbody_avg, relag, careerasp, TSO, avg_sexism, sexism_logel, howbody_1, howbody_2))

corr.test(select(smith_pair, physanx_A, a2pSOO_A, metaSOO_A, authenticity_A, howbody_avg_A, relag_A, careerasp_A, TSO_A, avg_sexism_A, sexism_logel_A))

corr.test(select(samegender_smith, a2pSOO_A, a2pSOO_P, metaSOO_A, metaSOO_P, howbody_avg_A))

corr.test(select(smith_pair, a2pSOO_A, a2pSOO_P, metaSOO_A, metaSOO_P, howbody_avg_A))
corr.test(select(smith_pair, authenticity_A, careerasp_A, relag_A))

#clark
corr.test(select(clark_final, physanx, a2pSOO, metaSOO, authenticity, howbody_avg, relag, careerasp, TSO, avg_sexism, sexism_logel, howbody_1, howbody_2))

corr.test(select(clark_pair, a2pSOO_A, a2pSOO_P, metaSOO_A, metaSOO_P, howbody_avg_A))
corr.test(select(clark_pair, authenticity_A, careerasp_A, relag_A))

#sexism histogram
#all gender
#ggplot(smith_final, aes(x=avg_sexism)) + geom_histogram(bins = 15)
#ggplot(clark_final, aes(x=avg_sexism)) + geom_histogram(bins=15)

#same

#A2p soo
#ggplot(smith_pair, aes(x=a2pSOO_A)) + geom_histogram(bins=15)
#ggplot(clark_pair, aes(x=a2pSOO_A)) + geom_histogram(bins=15)
#ggplot(samegender_smith, aes(x=a2pSOO_A)) + geom_histogram(bins=15)
#ggplot(samegender_clark, aes(x=a2pSOO_A)) + geom_histogram(bins=15)

#meta SOO
#ggplot(smith_pair, aes(x=metaSOO_A)) + geom_histogram(bins=15)
#ggplot(clark_pair, aes(x=metaSOO_A)) + geom_histogram(bins=15)
#ggplot(samegender_smith, aes(x=metaSOO_A)) + geom_histogram(bins=15)
#ggplot(samegender_clark, aes(x=metaSOO_A)) + geom_histogram(bins=15)

#TSO
# ggplot(smith_pair, aes(x=TSO_A)) + geom_histogram(bins=15)
# ggplot(clark_pair, aes(x=TSO_A)) + geom_histogram(bins=15)
# ggplot(samegender_smith, aes(x=TSO_A)) + geom_histogram(bins=15)
# ggplot(samegender_clark, aes(x=TSO_A)) + geom_histogram(bins=15)
# 
# #authenticity
# ggplot(smith_pair, aes(x=authenticity_A)) + geom_histogram(bins=10)
# ggplot(clark_pair, aes(x=authenticity_A)) + geom_histogram(bins=10)
# ggplot(samegender_smith, aes(x=TSO_A)) + geom_histogram(bins=15)
# ggplot(samegender_clark, aes(x=TSO_A)) + geom_histogram(bins=15)

#boxplots
#smith_howbody <- smith_final %>%
  #select(howbody_1, howbody_2, howbody_3, howbody_avg) %>%
  #gather(variable, value, howbody_1:howbody_avg)

#clark_howbody <- clark_final %>%
  #select(howbody_1, howbody_2, howbody_3, howbody_avg) %>%
  #gather(variable, value, howbody_1:howbody_avg)

#smith_howbody$value <- as.numeric(smith_howbody$value)
#clark_howbody$value <- as.numeric(clark_howbody$value)

# ggplot(smith_howbody, aes(x=variable, y=value)) + geom_boxplot()
# ggplot(clark_howbody, aes(x=variable, y=value)) + geom_boxplot()
# 
# ggplot(samegender_clark, aes(x=gender_A, y=a2pSOO_A)) + geom_boxplot()
# ggplot(samegender_clark, aes(x=gender_A, y=TSO_A)) + geom_boxplot()
# ggplot(clark_pair, aes(x=gender_A, y=TSO_A)) + geom_boxplot()

#howbody
corr.test(select(smith_final, howbody_1, howbody_2, howbody_3))
corr.test(select(clark_final, howbody_1, howbody_2, howbody_3))
psych::alpha(select(smith_final, howbody_1, howbody_2, howbody_3))
psych::alpha(select(clark_final, howbody_1, howbody_2, howbody_3))

# ggplot(smith_pair, aes(x=howbody_avg_A)) + geom_histogram(bins=15)
# ggplot(clark_pair, aes(x=howbody_avg_A)) + geom_histogram(bins=15)
# ggplot(samegender_smith, aes(x=howbody_avg_A)) + geom_histogram(bins=15) 
# ggplot(samegender_clark, aes(x=howbody_avg_A)) + geom_histogram(bins=15)

psych::alpha(select(smith_final, howbody_1, howbody_2, howbody_3))

test <- clark_measures[,140:150]
```

MODELING THE APIM
```{r}
#merging clark and smith same sex data
samegender_smith <- samegender_smith %>%
  mutate(smith_or_clark = "Smith",
         dyadID = dyadID + 100)

samegender_clark <- samegender_clark %>%
  mutate(smith_or_clark = "Clark")

clarksmith_same <- merge(x=samegender_smith, y=samegender_clark, all.x=TRUE, all.y=TRUE)

clarksmith_women <- clarksmith_same %>%
  filter(gender_A==2)

#indistinguishable dyads

apim <- gls(howbody_avg_A ~ a2pSOO_A + a2pSOO_P + TSO_A + TSO_A*a2pSOO_P,
            data=clarksmith_same,
            correlation=corCompSymm(form=~1|dyadID),
            na.action = na.omit)
summary(apim)

apim_in_empty <- gls(satisfaction_A ~ 1,
                     data = clarksmith_same,
                     correlation = corCompSymm(form=~1|dyadID),
                     na.action = na.omit)
summary(apim_in_empty)

r2 <- 1-(apim$sigma^2/apim_in_empty$sigma^2)
r2
```

##Descriptive stats
```{r}
favstats(~TSO_A, data=clarksmith_women)
favstats(~a2pSOO_A, data=clarksmith_women)
favstats(~howbody_avg_A, data=clarksmith_women)
favstats(~relag_A, data=clarksmith_women)
favstats(~careerasp_A, data=clarksmith_women)
favstats(~authenticity_A, data=clarksmith_women)
favstats(~trigram_score_A, data=clarksmith_women)

corr.test(select(clarksmith_women, TSO_A, authenticity_A, a2pSOO_A, howbody_avg_A, relag_A, careerasp_A,  trigram_score_A, TSO_P, authenticity_P, a2pSOO_P, howbody_avg_P, relag_P, careerasp_P, trigram_score_P))
```

##Participant info
```{r}
clark_final$age <- as.numeric(clark_final$age)
smith_final$age <- as.numeric(smith_final$age)
favstats(~age, data=clark_final)
favstats(~age, data=smith_final)

favstats(~age_A, data=clarksmith_women)
tally(clarksmith_women$race_A==clarksmith_women$race_P)

psych::alpha(select(clarksmith_women, a2pSOO_A, a2pSOO_P))

tally(~race_A, data=clarksmith_women)
tally(~sexor_A, data=clarksmith_women)
tally(~gender_A, data=clarksmith_women)

table(clarksmith_women$race_A, clarksmith_women$race_P)

test <- clarksmith_women %>%
  filter(race_A==3, race_A!=race_P)

tally(~gender, data=clark)
tally(~gender, data=smith)

```

```{r}
#Running direct correlations
SSO_authenticity <- gls(howbody_avg_A~authenticity_A, data=clarksmith_women)
auth_RAT <- gls(authenticity_A~trigram_score_A, data=clarksmith_women, na.action=na.omit)
auth_ASP <- gls(authenticity_A ~ careerasp_A, data=clarksmith_women)
auth_relag <- gls(authenticity_A ~ relag_A, data=clarksmith_women)

SSO_RAT <- gls(howbody_avg_A~trigram_score_A, data=clarksmith_women, na.action=na.omit)

summary(SSO_authenticity)
summary(auth_ASP)
summary(auth_relag)
summary(auth_RAT)
summary(SSO_RAT)
```

PATH model
```{r}
dyad_cswomen <- clarksmith_women %>%
  mutate(TSO_P_SOO_A = TSO_P*a2pSOO_A,
         TSO_A_SOO_P = TSO_A*a2pSOO_P) %>%
  filter(partnum_A==1)
```

```{r}
model <- '#####regressions 
howbody_avg_A ~ t*TSO_A + a*a2pSOO_A + b*a2pSOO_P
howbody_avg_P ~ t*TSO_P + a*a2pSOO_P + b*a2pSOO_A
authenticity_A ~ d*howbody_avg_A
authenticity_P ~ d*howbody_avg_P
trigram_score_A ~ e1*authenticity_A
careerasp_A ~ e2*authenticity_A
relag_A ~ e3*authenticity_A
trigram_score_P ~ e1*authenticity_P
careerasp_P ~ e2*authenticity_P
relag_P ~ e3*authenticity_P
######covariances
#endogenous
howbody_avg_A ~~ howbody_avg_P
authenticity_A ~~ authenticity_P
trigram_score_A ~~ trigram_score_P
careerasp_A ~~ careerasp_P
relag_A ~~ relag_P
trigram_score_A ~~ 0*careerasp_A
trigram_score_A ~~ 0*relag_A
trigram_score_A ~~ 0*careerasp_P
trigram_score_A ~~ 0*relag_P
careerasp_A ~~ 0*relag_A
careerasp_A ~~ 0*trigram_score_P
careerasp_A ~~ 0*relag_P
relag_A ~~ 0*trigram_score_P
relag_A ~~ 0*careerasp_P
trigram_score_P ~~ 0*careerasp_P
trigram_score_P ~~ 0*relag_P
careerasp_P ~~ 0*relag_P
#####variances
howbody_avg_A ~~ v1*howbody_avg_A
howbody_avg_P ~~ v1*howbody_avg_P
TSO_A ~~ v2*TSO_A
TSO_P ~~ v2*TSO_P
a2pSOO_A ~~ v3*a2pSOO_A
a2pSOO_P ~~ v3*a2pSOO_P
authenticity_A ~~ v4*authenticity_A
authenticity_P ~~ v4*authenticity_P
##endogenous intercepts
howbody_avg_A ~ i*1
howbody_avg_P ~ i*1
authenticity_A ~ i2*1
authenticity_P ~ i2*1
trigram_score_A ~ i3*1
trigram_score_P ~ i4*1
careerasp_A ~ i5*1
careerasp_P ~ i5*1
relag_A ~ i6*1
relag_P ~ i6*1
##exogenous means
TSO_A ~ m*1
TSO_P ~ m*1
a2pSOO_A ~ m2*1
a2pSOO_P ~ m2*1
TSO_P_SOO_A ~ m3*1
TSO_A_SOO_P ~ m3*1
## exogenous covariances
TSO_A_SOO_P ~~ c1*TSO_A
TSO_P_SOO_A ~~ c1*TSO_P
TSO_A_SOO_P ~~ c2*a2pSOO_P
TSO_P_SOO_A ~~ c2*a2pSOO_A
TSO_A ~~ c3*a2pSOO_P
TSO_P ~~ c3*a2pSOO_A
TSO_A ~~ c4*a2pSOO_A
TSO_P ~~ c4*a2pSOO_P
'

fit <- sem(model, data=dyad_cswomen)
summary(fit)
```