### 2019-04-01 Final Manuscript.Rmd
Adding alphas to manuscript, fixing manuscript data file, other stuff!!

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Kat Kyuchukova"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Asha Hinson"
    affiliation   : "1"
  - name          : "Randi L. Garcia"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Smith College"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library("papaja")
library(nlme)
library(haven)
library(devtools)
library(psych)
library(mosaic)
library(tidyr)
library(dplyr)
library(irr)

clark_measures <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Measures_1.sav")
smith <- read_spss("Raw data/Smith College Relationships Data/College_Relationship_Dyadic_September_14_2018_12_17.sav")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r, include=FALSE}
#######Alphas

###SOO

#Clark
#SOO
clark_extsoo_alpha <- psych::alpha(select(clark_measures, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10))$total[2]
clark_intsoo_alpha <- psych::alpha(select(clark_measures, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12))
#Smith
smith_extsoo_alpha <- psych::alpha(select(smith, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10))$total[2]
smith_intsoo_alpha <- psych::alpha(select(smith, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12))$total[2]

##Interaction Authenticity
#Clark
clark_intauth_alpha <- psych::alpha(select(clark_measures, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))$total[2]
#Smith
smith_intauth_alpha <- psych::alpha(select(smith, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))$total[2]

##SSO
#Clark
clark_howbody_alpha <- psych::alpha(select(clark_measures, howbody_1, howbody_2))$total[2]
#Smith
smith_howbody_alpha <- psych::alpha(select(smith, howbody_1, howbody_2))$total[2]

##Relationship Agency
#Clark
clark_relag_alpha <- psych::alpha(select(clark_measures, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))$total[2]
smith_relag_alpha <- psych::alpha(select(smith, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))$total[2]

##Career Aspirations
#Clark
clark_measures <- clark_measures %>% #reverse code certain career aspiration variables
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

clark_careerasp_alpha <- psych::alpha(select(clark_measures, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))$total[2]

#Smith
smith <- smith %>% #reverse code certain career aspiration variables
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

smith_careerasp_alpha <- psych::alpha(select(smith, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))$total[2]

###TSO
#Clark--external and internal
clark_funcTSO_alpha <- psych::alpha(select(clark_measures,tso_1, tso_2, tso_4, tso_7, tso_9))$total[2]
clark_physTSO_alpha <- psych::alpha(select(clark_measures,tso_3, tso_5, tso_6, tso_8, tso_10))$total[2]
#Smith--external and internal
smith_funcTSO_alpha <- psych::alpha(select(smith,tso_1, tso_2, tso_4, tso_7, tso_9))$total[2]
smith_physTSO_alpha <- psych::alpha(select(smith,tso_3, tso_5, tso_6, tso_8, tso_10))$total[2]
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants
The data collected in this study sought to replicate Garcia et al. (2016), which sought to measure objectification in previously unacquainted male-female pairs. This data was given to us by Dr. Garcia, who collected data from both Smith College and Clark University. The data was ultimately merged and filtered to only include only self-identified female same-sex pairs from both schools. The two samples are often referred to below as Sample 1 (data from Clark University) and Sample 2 (data from Smith College). These participants were mostly first-year college students, with an average age of 18.85 years (SD=1.04). The sample was 48% White/European American, 9% Black/African-American, 28% Asian/Pacific Islander, 9% Latinx, and 5% mixed-race. There were eight White/White pairs and four same race racial minority pairs, for a total of twelve same-race pairs. The remaining twenty were mixed race pairs, of which fifteen were White/racial minority pairings and five were cross-racial minority group pairs. 64% of the sample identified as heterosexual, and 23% identified as gay, lesbian or bisexual. The data is organized in a pairwise fashion such that each row corresponds to one participant in the study, and includes data for themselves and their partner. We would want to generalize these findings to general interactions between women.

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup