### 2019-04-01 Variable creation.Rmd

---
title: "Variable creation ObjectInAction"
author: "Kat Kyuchukova"
date: "11/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library(nlme)
library(haven)
library(devtools)
library(psych)
library(mosaic)
library(tidyr)
library(dplyr)
library(irr)
select <- dplyr::select
```

Load data
```{r}
#clark
clark_measures <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Measures_1.sav")
clark_pairwise <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Pairwise_1.sav")

#smith
smith <- read_spss("Raw data/Smith College Relationships Data/College_Relationship_Dyadic_September_14_2018_12_17.sav")

clark_sentences <- read.csv("Raw data/Clark College Relationships Data/sexist sentences/coded_sentences.csv")

#trigrams
trigram_check <- read.csv("Raw data/Clark College Relationships Data/trigrams/AMELIA_check_trigram_grading.csv")
```

Clark Scale Scores--creating scale scores for the Clark data.
```{r}
#SOO -- subtract internal SOO from external SOO
clark_measures$extSOO <- rowMeans(dplyr::select(clark_measures, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE) #get average

clark_measures$intSOO <- rowMeans(dplyr::select(clark_measures, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE) #get average

clark_measures$a2pSOO <- clark_measures$extSOO - clark_measures$intSOO #subtract internal from external

#Meta SOO--subtract internal from external
clark_measures$extmetaSOO <- rowMeans(dplyr::select(clark_measures, metaint_1, metaint_2, metaint_5, metaint_6, metaint_10), na.rm=TRUE)

clark_measures$intmetaSOO <- rowMeans(dplyr::select(clark_measures, metaint_3, metaint_4, metaint_7, metaint_8, metaint_9, metaint_11, metaint_12), na.rm=TRUE)

clark_measures$metaSOO <- clark_measures$extmetaSOO - clark_measures$intmetaSOO

#Interaction Authenticity
clark_measures$authenticity <- rowMeans(dplyr::select(clark_measures, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

#SSO--average of how body 1 and how body 2
clark_measures <- clark_measures %>%
  mutate(howbody_avg = (howbody_1+howbody_2)/2)

#Relationship Agency
clark_measures$relag <- rowMeans(dplyr::select(clark_measures, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

#Career Aspirations--reverse code 4, 7, and 10 and average with the rest
clark_measures <- clark_measures %>% #reverse code certain career aspiration variables
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

clark_measures$careerasp <- rowMeans(dplyr::select(clark_measures, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

#TSO--subtract internal from external, multiply by -1
clark_measures$funcTSO <- rowMeans(dplyr::select(clark_measures,tso_1, tso_2, tso_4, tso_7, tso_9))
clark_measures$physTSO <- rowMeans(dplyr::select(clark_measures, tso_3, tso_5, tso_6, tso_8, tso_10))

clark_measures$TSO <- clark_measures$physTSO - clark_measures$funcTSO
clark_measures$TSO <- clark_measures$TSO*(-1)
```

SMITH SCALE SCORES
```{r}
smith <- smith %>%
  arrange(dyadID) %>%
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999) #filter out dyads with no data

#SOO
smith$extSOO <- rowMeans(dplyr::select(smith, otherint_1, otherint_2, otherint_5, otherint_6, otherint_10), na.rm=TRUE)

smith$intSOO <- rowMeans(dplyr::select(smith, otherint_3, otherint_4, otherint_7, otherint_8, otherint_9, otherint_11, otherint_12), na.rm=TRUE)

smith$a2pSOO <- smith$extSOO - smith$intSOO

#Meta SOO
smith$extmetaSOO <- rowMeans(dplyr::select(smith, metaint_1, metaint_2, metaint_5, metaint_6, metaint_10), na.rm=TRUE)

smith$intmetaSOO <- rowMeans(dplyr::select(smith, metaint_3, metaint_4, metaint_7, metaint_8, metaint_9, metaint_11, metaint_12), na.rm=TRUE)

smith$metaSOO <- smith$extmetaSOO - smith$intmetaSOO

#Interaction Authenticity
smith$authenticity <- rowMeans(dplyr::select(smith, inter_1, inter_3, inter_5, inter_6, inter_7, inter_9, inter_10, interaux_5, interaux_6))

#SSO
smith <- smith %>%
  mutate(howbody_avg = (howbody_1+howbody_2)/2)

#Relationship Agency
smith$relag <- rowMeans(dplyr::select(smith, relag_1, relag_2, relag_3, relag_4, relag_5, relag_6, relag_7, relag_8, relag_9))

#Career Aspirations
smith <- smith %>%
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

smith$careerasp <- rowMeans(dplyr::select(smith, careerasp_1, careerasp_2, careerasp_3, careerasp4.r, careerasp_5, careerasp_6, careerasp7.r, careerasp_8, careerasp_9, careerasp10.r))

#TSO
smith$funcTSO <- rowMeans(dplyr::select(smith,tso_1, tso_2, tso_4, tso_7, tso_9))
smith$physTSO <- rowMeans(dplyr::select(smith, tso_3, tso_5, tso_6, tso_8, tso_10))

smith$TSO <- smith$physTSO-smith$funcTSO
smith$TSO <- smith$TSO*-1
```

Trigrams
```{r}
###Clark
#Eliminate unnecessary columns
trigram_check <- trigram_check[,1:8]

#correct dummy variable to account for the correct trigrams that excel missed
trigram_check <- trigram_check %>%
  mutate(trigram_correct = ifelse(accurate.=="yes", Correct...1.0., 1))

#create a correct trigram score for each participant
trigram_clark <- trigram_check %>%
  group_by(partID) %>%
  summarise(trigram_score = sum(trigram_correct)) %>%
  dplyr::select(partID, trigram_score)

#merge trigram data with big clark data file
clark <- merge(x=clark_measures, y=trigram_clark, by="partID", all.x=TRUE)
```

```{r}
###SMITH
#select only trigram related variables
smith_trigrams <- smith %>%
  dplyr::select(partID, dyadID, rat1_1.0, rat2_1.0,rat3_1.0, rat4_1.0,rat5_1.0, rat6_1.0, rat7_1.0, rat8_1.0, rat9_1.0, rat10_1) %>%
  arrange(dyadID, partID)

smith_trigrams_long <- gather(smith_trigrams, trigram, response, rat1_1.0:rat10_1)

#change trigram from generic name to the specific trigram
smith_trigrams_long <- smith_trigrams_long %>%
  mutate(trigram = ifelse(trigram=="rat1_1.0", "QUACK:POND:WADDLE",
                          ifelse(trigram=="rat2_1.0", "ATHLETES:WEB:RABBIT",
                                 ifelse(trigram=="rat3_1.0", "HONEY:SWARM:STING",
                                        ifelse(trigram=="rat4_1.0", "MATCHES:SMOKE:BUSH",
                                               ifelse(trigram=="rat5_1.0", "CHOCOLATE:FORTUNE:TIN",
                                                      ifelse(trigram=="rat6_1.0", "CURIOSITY:NAP:WHISKERS",
                                                             ifelse(trigram=="rat7_1.0", "STALK:TRAINER:KING",
                                                                    ifelse(trigram=="rat8_1.0", "SURPRISE:LINE:BIRTHDAY",
                                                                           ifelse(trigram=="rat9_1.0", "RED:GO:CAR", 
                                                                                  ifelse(trigram=="rat10_1", "MOUSE : SHARP : BLUE", NA)))))))))))

#create a correct response column
smith_trigrams_long <- smith_trigrams_long %>%
  mutate(correct_response=ifelse(trigram=="QUACK:POND:WADDLE", "Duck",
                          ifelse(trigram=="ATHLETES:WEB:RABBIT", "Foot",
                                 ifelse(trigram=="HONEY:SWARM:STING","Bee",
                                        ifelse(trigram=="MATCHES:SMOKE:BUSH", "Fire", 
                                               ifelse(trigram=="CHOCOLATE:FORTUNE:TIN", "Cookie",
                                                      ifelse(trigram=="CURIOSITY:NAP:WHISKERS","Cat",
                                                             ifelse(trigram=="STALK:TRAINER:KING","Lion",
                                                                    ifelse(trigram=="SURPRISE:LINE:BIRTHDAY", "Party",
                                                                           ifelse(trigram=="RED:GO:CAR", "Stop",
                                                                                  ifelse(trigram=="MOUSE : SHARP : BLUE", "Cheese", NA))))))))))) %>%
  arrange(partID, dyadID)

smith_trigrams_long <- smith_trigrams_long[,c(1,2,3,5,4)]

##Write out file to grade
#write.csv(smith_trigrams_long, "Raw data/Smith College Relationships Data/trigrams_to_grade.csv", row.names=FALSE)

#Read in graded trigram file
smith_trigrams_graded <- read.csv("Raw data/Smith College Relationships Data/trigrams_graded.csv")

#Summarize a final correct trigram
smith_triscore <- smith_trigrams_graded %>%
  group_by(partID, dyadID) %>%
  summarise(trigram_score=sum(final_correct))%>%
  dplyr::select(partID, dyadID, trigram_score)

#Merge with big smith data
smith <- merge(x=smith, y=smith_triscore, x.all=TRUE)
```

sexist sentences
```{r}
#Calculate ICC score to see if we can average
clark_sentences <- clark_sentences[,1:8]

clark_sentences <- clark_sentences %>%
  filter(!is.na(dyadID))

clark_sentences <- na.omit(clark_sentences)
clark_icc <- clark_sentences %>%
  select(Hannah, Jennie, Amelia)

icc(clark_icc, model="twoway", unit="average")

#Remove sentence 4
clark_sentences <- clark_sentences %>%
  filter(Sentence!=4)

#ICC is good enough to get an average

#Total sentence average
clark_sentence_avg <- clark_sentences %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

#sentence average with just the logel sentences
clark_logel <- clark_sentences %>%
  filter(Sentence==3|Sentence==5|Sentence==6) %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

#Merge together 
clark_sexism <- merge(x=clark, y=clark_sentence_avg, by="partID", all.x=TRUE)
clark_final <- merge (x=clark_sexism, y=clark_logel, by="partID", all.x=TRUE)
```

```{r}
#Smith

#Creating sexist sentence file
smith$dyadID <- as.numeric(smith$dyadID)
smith$partID <- as.numeric(smith$partID)
smith_sentences_long <- smith %>%
  #remove dyads with no data
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999) %>%
  #select just sexism related variables
  select(partID, dyadID, sexismpre_1:sexismpre_8) %>%
  arrange(dyadID)

#make data long
smith_sentences_long <- gather(smith_sentences_long, sentence, target, sexismpre_1:sexismpre_8)

#create a column with the actual sentence as well as a sentence number column
smith_sentences_long <- smith_sentences_long %>%
  mutate(stem=ifelse(sentence=="sexismpre_1", "Linda swatted at the flies", 
                     ifelse(sentence=="sexismpre_2", "Helen crammed for the test",
                            ifelse(sentence=="sexismpre_3", "Shirley asked for help getting home",
                                   ifelse(sentence=="sexismpre_4", "Sam read the newspaper",
                                          ifelse(sentence=="sexismpre_5", "Jenny went home to cook dinner",
                                                 ifelse(sentence=="sexismpre_6", "Katherine baby-sat the neighbor's kids",
                                                        ifelse(sentence=="sexismpre_7", "Tom bottle fed the baby",
                                                               ifelse(sentence=="sexismpre_8", "Bert changed the oil", NA))))))))) %>%
    mutate(sentence=ifelse(sentence=="sexismpre_1", 1, 
                     ifelse(sentence=="sexismpre_2", 2,
                            ifelse(sentence=="sexismpre_3", 3,
                                   ifelse(sentence=="sexismpre_4", 4,
                                          ifelse(sentence=="sexismpre_5", 5,
                                                 ifelse(sentence=="sexismpre_6", 6,
                                                        ifelse(sentence=="sexismpre_7", 7,
                                                               ifelse(sentence=="sexismpre_8", 8, NA)))))))))

#change order of sentences
smith_sentences_long <- smith_sentences_long[,c(1,2,3,5,4)]

#remove rows with no target data
smith_sentences_long <- smith_sentences_long %>%
  filter(target!="") %>%
  arrange(dyadID, partID)

#Write out csv
#write.csv(smith_sentences_long, "Raw data/Smith College Relationships Data/sexist_sentences_for_coding.csv", row.names = FALSE)

#Read in coded sentences, code correlations
kat_codes <- read.csv("Raw data/Smith College Relationships Data/kat_recode.csv")

xinyu_codes <- read.csv("Raw data/Smith College Relationships Data/xinyu_recode.csv")

coded_sentences <- merge(x=kat_codes, y=xinyu_codes, all.x = TRUE)

smith_icc <- coded_sentences %>%
  #remove sentence 4
  filter(sentence!=4)

icc(smith_icc, model="twoway", unit="average")

#making average file
#all sentenes (minus 4)
smith_sentence_avg <- coded_sentences %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

#just logel sentences
smith_logel <- coded_sentences %>%
  filter(sentence==3|sentence==5|sentence==6) %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

#merge data
smith_sexism <- merge(x=smith, y=smith_sentence_avg, by="partID", all.x=TRUE)
smith_final <- merge (x=smith_sexism, y=smith_logel, by="partID", all.x=TRUE)
```

Write out cleaned data!
```{r}
write.csv(smith_final, "ObjectInAction/Cleaned data/smith_final.csv" )
write.csv(clark_final, "ObjectInAction/Cleaned data/clark_final.csv")
```