### 2020-01-23 i-sat_fit_statistics.Rmd
some random files I had

```{r}
chi_isat <- 89.135
df_isat <- 75

#this is definitely right  
chi_mod <- 120.502
df_mod <- 112

chi_null <- 218.844
df_null <- 129
```

adjust

```{r}
chi_mod_adj <- chi_mod - chi_isat
df_mod_adj <- df_mod - df_isat

chi_null_adj <- chi_null - chi_isat
df_null_adj <- df_null - df_isat
```

Chi-square test for model fit

```{r}
pchisq(chi_mod_adj, df_mod_adj, lower.tail = FALSE)
```

Fit statistics

```{r}
RMSEA <- sqrt(((chi_mod_adj/df_mod_adj)-1)/(32-1))
chi_mod_adj/df_mod_adj-1
sqrt(-.15/31)

#Trying to take square root of negative number...so zero. Because df is smaller than chi-sqaure. 
RMSEA <- 0 
```

```{r}
CFI <- 1-((chi_mod_adj-df_mod_adj)/(chi_null_adj-df_null_adj))
CFI
```