### 2020-07-06 Variable creation.Rmd
Cleaned file i-o

---
title: "Variable creation ObjectInAction"
author: "Kat Kyuchukova"
date: "11/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library(nlme)
library(haven)
library(devtools)
library(psych)
library(mosaic)
library(tidyr)
library(irr)
library(dplyr)
```

#Load data
```{r}
#clark
clark_measures <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Measures_1.sav")
clark_pairwise <- read_spss("Raw data/Clark College Relationships Data/College_Relationship_Pairwise_1.sav")

#smith
smith <- read_spss("Raw data/Smith College Relationships Data/College_Relationship_Dyadic_September_14_2018_12_17.sav")

clark_sentences <- read.csv("Raw data/Clark College Relationships Data/sexist sentences/coded_sentences.csv")

#trigrams
trigram_check <- read.csv("Raw data/Clark College Relationships Data/trigrams/AMELIA_check_trigram_grading.csv")
```

#Scale Scores

Creating scale scores for the Clark data.

```{r}
#SOO -- subtract internal SOO from external SOO
clark_measures$extSOO <- rowMeans(select(clark_measures, otherint_1, otherint_2, 
                                         otherint_5, otherint_6, otherint_10), na.rm=TRUE) #get average

clark_measures$intSOO <- rowMeans(select(clark_measures, otherint_3, otherint_4, 
                                         otherint_7, otherint_8, otherint_9, 
                                         otherint_11, otherint_12), na.rm=TRUE) #get average

clark_measures$a2pSOO <- clark_measures$extSOO - clark_measures$intSOO #subtract internal from external

#Meta SOO--subtract internal from external
clark_measures$extmetaSOO <- rowMeans(select(clark_measures, metaint_1, metaint_2, 
                                             metaint_5, metaint_6, metaint_10), na.rm=TRUE)

clark_measures$intmetaSOO <- rowMeans(select(clark_measures, metaint_3, metaint_4, 
                                             metaint_7, metaint_8, metaint_9, 
                                             metaint_11, metaint_12), na.rm=TRUE)

clark_measures$metaSOO <- clark_measures$extmetaSOO - clark_measures$intmetaSOO

#Interaction Authenticity
clark_measures$authenticity <- rowMeans(select(clark_measures, inter_1, inter_3, 
                                               inter_5, inter_6, inter_7, inter_9, 
                                               inter_10, interaux_5, interaux_6), na.rm=TRUE)

#SSO--average of how body 1 and how body 2
clark_measures$howbody_avg <- rowMeans(select(clark_measures, howbody_1, howbody_2), na.rm=TRUE)

#Relationship Agency
#Reverse code relationship agency 9
clark_measures <- clark_measures %>%
  mutate(relag_9.r=(min(relag_9, na.rm=TRUE) + max(relag_9, na.rm=TRUE) - relag_9))

clark_measures$relag <- rowMeans(select(clark_measures, relag_1, relag_2, 
                                        relag_3, relag_4, relag_5, relag_6, 
                                        relag_7, relag_8, relag_9.r), na.rm=TRUE)

#Career Aspirations--reverse code 4, 7, and 10 and average with the rest
clark_measures <- clark_measures %>% #reverse code certain career aspiration variables
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

clark_measures$careerasp <- rowMeans(select(clark_measures, careerasp_1, careerasp_2, 
                                            careerasp_3, careerasp4.r, careerasp_5, 
                                            careerasp_6, careerasp7.r, careerasp_8, 
                                            careerasp_9, careerasp10.r), na.rm=TRUE)

#TSO--subtract internal from external, multiply by -1
clark_measures$funcTSO <- rowMeans(select(clark_measures,tso_1, tso_2, tso_4, tso_7, tso_9), na.rm=TRUE)
clark_measures$physTSO <- rowMeans(select(clark_measures, tso_3, tso_5, tso_6, tso_8, tso_10), na.rm=TRUE)

clark_measures$TSO <- (clark_measures$physTSO - clark_measures$funcTSO)*(-1)
```

Creating scale scores for the Smith data.

```{r}
smith <- smith %>%
  arrange(dyadID) %>%
  filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, 
         dyadID!=999, dyadID!=999999, dyadID!=9999) #filter out dyads with no data
```

```{r}
#SOO
smith$extSOO <- rowMeans(select(smith, otherint_1, otherint_2, 
                                otherint_5, otherint_6, otherint_10), na.rm=TRUE)

smith$intSOO <- rowMeans(select(smith, otherint_3, otherint_4, 
                                otherint_7, otherint_8, otherint_9, 
                                otherint_11, otherint_12), na.rm=TRUE)

smith$a2pSOO <- smith$extSOO - smith$intSOO

#Meta SOO
smith$extmetaSOO <- rowMeans(select(smith, metaint_1, metaint_2, 
                                    metaint_5, metaint_6, metaint_10), na.rm=TRUE)

smith$intmetaSOO <- rowMeans(select(smith, metaint_3, metaint_4, 
                                    metaint_7, metaint_8, metaint_9, 
                                    metaint_11, metaint_12), na.rm=TRUE)

smith$metaSOO <- smith$extmetaSOO - smith$intmetaSOO

#Interaction Authenticity
smith$authenticity <- rowMeans(select(smith, inter_1, inter_3, inter_5, 
                                      inter_6, inter_7, inter_9, inter_10, 
                                      interaux_5, interaux_6), na.rm=TRUE)

#SSO
smith$howbody_avg <- rowMeans(select(smith, howbody_1, howbody_2), na.rm=TRUE)

#Relationship Agency

smith <- smith %>%
  mutate(relag_9.r=(min(relag_9, na.rm=TRUE) + max(relag_9, na.rm=TRUE) - relag_9))

smith$relag <- rowMeans(select(smith, relag_1, relag_2, relag_3, 
                               relag_4, relag_5, relag_6, relag_7, 
                               relag_8, relag_9.r), na.rm=TRUE)

#Career Aspirations
smith <- smith %>%
  mutate(careerasp4.r = (min(careerasp_4, na.rm = TRUE) + max(careerasp_4, na.rm = TRUE) - careerasp_4),
         careerasp7.r = (min(careerasp_7, na.rm = TRUE) + max(careerasp_7, na.rm = TRUE) - careerasp_7),
         careerasp10.r = (min(careerasp_10, na.rm = TRUE) + max(careerasp_10, na.rm = TRUE) - careerasp_10))

smith$careerasp <- rowMeans(select(smith, careerasp_1, careerasp_2, 
                                   careerasp_3, careerasp4.r, careerasp_5, 
                                   careerasp_6, careerasp7.r, careerasp_8, 
                                   careerasp_9, careerasp10.r), na.rm=TRUE)

#TSO
smith$funcTSO <- rowMeans(select(smith,tso_1, tso_2, tso_4, tso_7, tso_9), na.rm=TRUE)
smith$physTSO <- rowMeans(select(smith, tso_3, tso_5, tso_6, tso_8, tso_10), na.rm=TRUE)

smith$TSO <- (smith$physTSO - smith$funcTSO)*(-1)
```

Step not needed...processing continues
```{r}
#Write out csv with created variables
#write.csv(clark_measures, "Cleaned data/clark_measures.csv")
#write.csv(smith, "Cleaned data/smith.csv")
```

#Trigrams

Clark

```{r}
###Clark
#Eliminate unnecessary columns
trigram_check <- trigram_check[,1:8]

#correct dummy variable to account for the correct trigrams that excel missed
trigram_check <- trigram_check %>%
  mutate(trigram_correct = ifelse(accurate.=="yes", Correct...1.0., 1))

#create a correct trigram score for each participant
trigram_clark <- trigram_check %>%
  group_by(partID) %>%
  summarise(trigram_score = sum(trigram_correct)) %>%
  select(partID, trigram_score)

#merge trigram data with big clark data file
clark <- merge(x=clark_measures, y=trigram_clark, by="partID", all.x=TRUE)
```

Smith

```{r}
###SMITH

#comented code here used only ONCE to grade trigrams...graded trigrams file is read in at the bottom of this chunk

#select only trigram related variables
# smith_trigrams <- smith %>%
#   select(partID, dyadID, rat1_1.0, rat2_1.0,rat3_1.0, rat4_1.0,
#          rat5_1.0, rat6_1.0, rat7_1.0, rat8_1.0, rat9_1.0, rat10_1) %>%
#   arrange(dyadID, partID)
# 
# smith_trigrams_long <- gather(smith_trigrams, trigram, response, rat1_1.0:rat10_1)
# 
# #change trigram from generic name to the specific trigram
# smith_trigrams_long <- smith_trigrams_long %>%
#   mutate(trigram = ifelse(trigram=="rat1_1.0", 
#                           "QUACK:POND:WADDLE",
#                           ifelse(trigram=="rat2_1.0", 
#                                  "ATHLETES:WEB:RABBIT",
#                                  ifelse(trigram=="rat3_1.0", 
#                                         "HONEY:SWARM:STING",
#                                         ifelse(trigram=="rat4_1.0", 
#                                                "MATCHES:SMOKE:BUSH",
#                                                ifelse(trigram=="rat5_1.0", 
#                                                       "CHOCOLATE:FORTUNE:TIN",
#                                                       ifelse(trigram=="rat6_1.0", 
#                                                              "CURIOSITY:NAP:WHISKERS",
#                                                              ifelse(trigram=="rat7_1.0", 
#                                                                     "STALK:TRAINER:KING",
#                                                                     ifelse(trigram=="rat8_1.0", 
#                                                                            "SURPRISE:LINE:BIRTHDAY",
#                                                                            ifelse(trigram=="rat9_1.0", 
#                                                                                   "RED:GO:CAR", 
#                                                                                   ifelse(trigram=="rat10_1", 
#                                                                                          "MOUSE : SHARP : BLUE", 
#                                                                                          NA)))))))))))
# 
# #create a correct response column
# smith_trigrams_long <- smith_trigrams_long %>%
#   mutate(correct_response=ifelse(trigram=="QUACK:POND:WADDLE", 
#                                  "Duck",
#                           ifelse(trigram=="ATHLETES:WEB:RABBIT", 
#                                  "Foot",
#                                  ifelse(trigram=="HONEY:SWARM:STING",
#                                         "Bee",
#                                         ifelse(trigram=="MATCHES:SMOKE:BUSH", 
#                                                "Fire", 
#                                                ifelse(trigram=="CHOCOLATE:FORTUNE:TIN", 
#                                                       "Cookie",
#                                                       ifelse(trigram=="CURIOSITY:NAP:WHISKERS",
#                                                              "Cat",
#                                                              ifelse(trigram=="STALK:TRAINER:KING",
#                                                                     "Lion",
#                                                                     ifelse(trigram=="SURPRISE:LINE:BIRTHDAY", 
#                                                                            "Party",
#                                                                            ifelse(trigram=="RED:GO:CAR", 
#                                                                                   "Stop",
#                                                                                   ifelse(trigram=="MOUSE : SHARP : BLUE", 
#                                                                                          "Cheese", 
#                                                                                          NA))))))))))) %>%
#   arrange(partID, dyadID)
# 
# smith_trigrams_long <- smith_trigrams_long[,c(1,2,3,5,4)]

##Write out file to grade
#write.csv(smith_trigrams_long, "Raw data/Smith College Relationships Data/trigrams_to_grade.csv", row.names=FALSE)

#Read in graded trigram file
smith_trigrams_graded <- read.csv("Raw data/Smith College Relationships Data/trigrams_graded.csv")

#Summarize a final correct trigram
smith_triscore <- smith_trigrams_graded %>%
  group_by(partID, dyadID) %>%
  summarise(trigram_score=sum(final_correct))%>%
  select(partID, dyadID, trigram_score)

#Merge with big smith data
smith <- merge(x=smith, y=smith_triscore, x.all=TRUE)
```

#Sexist Sentences

Clark

```{r}
#Calculate ICC score to see if we can average
clark_sentences <- clark_sentences[,1:8]

clark_sentences <- clark_sentences %>%
  filter(!is.na(dyadID))

clark_sentences <- na.omit(clark_sentences)
clark_icc <- clark_sentences %>%
  select(Hannah, Jennie, Amelia)

icc(clark_icc, model="twoway", unit="average")

#Remove sentence 4
clark_sentences <- clark_sentences %>%
  filter(Sentence!=4)

#ICC is good enough to get an average

#Total sentence average
clark_sentence_avg <- clark_sentences %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

#sentence average with just the logel sentences
clark_logel <- clark_sentences %>%
  filter(Sentence==3|Sentence==5|Sentence==6) %>%
  mutate(sexism=(Hannah+Jennie+Amelia)/3) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

#Merge together 
clark_sexism <- merge(x=clark, y=clark_sentence_avg, by="partID", all.x=TRUE)
clark_final <- merge (x=clark_sexism, y=clark_logel, by="partID", all.x=TRUE)
```

Smith

```{r}
#Smith

#comented code here used only ONCE to code sentences...graded trigrams file is read in at the bottom of this chunk

#Creating sexist sentence file
# smith$dyadID <- as.numeric(smith$dyadID)
# smith$partID <- as.numeric(smith$partID)
# smith_sentences_long <- smith %>%
#   #remove dyads with no data
#   filter(dyadID!=16, !is.na(dyadID), dyadID!=798797, dyadID!=999, dyadID!=999999, dyadID!=9999) %>%
#   #select just sexism related variables
#   select(partID, dyadID, sexismpre_1:sexismpre_8) %>%
#   arrange(dyadID)
# 
# #make data long
# smith_sentences_long <- gather(smith_sentences_long, sentence, target, sexismpre_1:sexismpre_8)
# 
# #create a column with the actual sentence as well as a sentence number column
# smith_sentences_long <- smith_sentences_long %>%
#   mutate(stem=ifelse(sentence=="sexismpre_1", 
#                      "Linda swatted at the flies", 
#                      ifelse(sentence=="sexismpre_2", 
#                             "Helen crammed for the test",
#                             ifelse(sentence=="sexismpre_3", 
#                                    "Shirley asked for help getting home",
#                                    ifelse(sentence=="sexismpre_4", 
#                                           "Sam read the newspaper",
#                                           ifelse(sentence=="sexismpre_5", 
#                                                  "Jenny went home to cook dinner",
#                                                  ifelse(sentence=="sexismpre_6", 
#                                                         "Katherine baby-sat the neighbor's kids",
#                                                         ifelse(sentence=="sexismpre_7", 
#                                                                "Tom bottle fed the baby",
#                                                                ifelse(sentence=="sexismpre_8", 
#                                                                       "Bert changed the oil", NA))))))))) %>%
#     mutate(sentence=ifelse(sentence=="sexismpre_1", 1, 
#                      ifelse(sentence=="sexismpre_2", 2,
#                             ifelse(sentence=="sexismpre_3", 3,
#                                    ifelse(sentence=="sexismpre_4", 4,
#                                           ifelse(sentence=="sexismpre_5", 5,
#                                                  ifelse(sentence=="sexismpre_6", 6,
#                                                         ifelse(sentence=="sexismpre_7", 7,
#                                                                ifelse(sentence=="sexismpre_8", 8, NA)))))))))
# 
# #change order of sentences
# smith_sentences_long <- smith_sentences_long[,c(1,2,3,5,4)]
# 
# #remove rows with no target data
# smith_sentences_long <- smith_sentences_long %>%
#   filter(target!="") %>%
#   arrange(dyadID, partID)

#Write out csv
#write.csv(smith_sentences_long, "Raw data/Smith College Relationships Data/sexist_sentences_for_coding.csv", row.names = FALSE)

#Read in coded sentences, code correlations
kat_codes <- read.csv("Raw data/Smith College Relationships Data/kat_recode.csv")
xinyu_codes <- read.csv("Raw data/Smith College Relationships Data/xinyu_recode.csv")

coded_sentences <- merge(x=kat_codes, y=xinyu_codes, all.x = TRUE)

smith_icc <- coded_sentences %>%
  #remove sentence 4
  filter(sentence!=4)

icc(smith_icc, model="twoway", unit="average")

#making average file
#all sentenes (minus 4)
smith_sentence_avg <- coded_sentences %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(avg_sexism = mean(sexism)) %>%
  select(partID, avg_sexism)

#just logel sentences
smith_logel <- coded_sentences %>%
  filter(sentence==3|sentence==5|sentence==6) %>%
  mutate(sexism=(katscore2+Xinyu)/2) %>%
  group_by(partID) %>%
  summarise(sexism_logel = mean(sexism)) %>%
  select(partID, sexism_logel)

#merge data
smith_sexism <- merge(x=smith, y=smith_sentence_avg, by="partID", all.x=TRUE)
smith_final <- merge (x=smith_sexism, y=smith_logel, by="partID", all.x=TRUE)
```

Write out cleaned data!
```{r}
write.csv(smith_final, "Cleaned data/smith_final.csv" )
write.csv(clark_final, "Cleaned data/clark_final.csv")
```