### 2020-09-09 Final_Manuscript.tex
re-knit paper

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Objectification in Action: Self- and Other-Objectification in Same-gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Laneé Jung},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Objectification in Action: Self- and Other-Objectification in
Same-gender Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1}, Asha Hinson\textsuperscript{1}, \& Laneé
Jung\textsuperscript{1}}
    \date{}

\shorttitle{Objectification in Action}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Smith College\\\textsuperscript{2} Program in Statistical and Data Sciences, Smith College}
\keywords{keywords}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\authornote{

Correspondence concerning this article should be addressed to Randi L.
Garcia, 415 Bass Hall, Smith College, Northampton, MA 01060. E-mail:
\href{mailto:rgarcia@smith.edu}{\nolinkurl{rgarcia@smith.edu}}}

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for women within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between objectification and self-objectification within interactions
between female pairs. Women were brought into the laboratory and
interacted in same-sex dyads. Dyadic analysis was utilized to detect
whether partners' objectification of each other affected state
self-objectification, as well as the resulting feelings of comfort and
authenticity during the interaction. After the interaction, participants
completed a questionnaire measuring, among other things, cognitive
performance, career aspirations, and relationship agency. Results
revealed a significant positive effect of being objectified by a female
partner on women's self-objectification. There was no significant
relationship between self-objectification and authenticity. Further,
although there were significantly negative effects on career aspirations
and relationship agency resulting from a lack of relationship
authenticity, there was no evidence that this is due to feelings of
sexual objectification. The significant partner effect of
objectification on actor self-objectification suggests that women being
objectified by other women may result in feelings of
self-objectification, and such research has powerful implications for
the ways that women interact in both sexual and non-sexual settings.

}

\begin{document}
\maketitle

Roberts and Fredrickson (1997) Objectification Theory, suggests that, in
addition to being steeped in a culture that objectifies women, women are
objectified in actual interpersonal encounters. The negative effects of
this interpersonal objectification are theorized be the strongest when
it is a man doing the objectifying (Gervais, Vescio, \& Allen, 2011),
but what are the consequences when a woman objectifies another woman in
an interaction? Psychological researchers studying the sexual
objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
self-objectification in the target as a proximal consequence (Garcia,
Earnshaw, \& Quinn, 2016, Strelan and Pagoudis (2018), Riemer, Sáez,
Brock, and Gervais (2020)). For example, in a face-to-face mixed-gender
dyadic study, Garcia et al. (2016) found that men's reported
objectification of their female interaction partner was associated with
increased self-objectification reported by their partner. They did not
include same-gender, woman-woman interacting pairs. Gervais, Sáez,
Riemer, and Klein (2020) have reviewed the literature on interpersonal
self-objectification and has organized the process in a theoretical
model called the Social Interaction Model of Objectification (SIMO), but
this model is currently only relevant for mixed-gender encounters.
Although there is ample evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), there are currently no
studies investigating the process of interpersonal objectification in
same-gender interactions. Does objectification by a female interaction
partner have the same downstream negative consequences as being
objectified by a male interaction partner? The current dyadic study
addresses this omission using interacting pairs where both members
identify as women.

\subsection{Self-Objectification}\label{self-objectification}

Self-objectification is a psychological process that provides a
cognitive mechanism to translate the experiences of interpersonal
objectification (Loughnan et al., 2015a) to negative mental health
outcomes (e.g., anxiety, self-esteem, and cognitive performance)
(Calogero, Tantleff-Dunn, \& Thompson, 2011; Moradi \& Huang, 2008).
Calogero et al. (2011) proposes that self-objectification can be
conceptualized as a learned trait, or \emph{trait} self-objectification
(TSO). Furthermore, it can also be elicited momentarily, through the
media, for example, with sexualized images in movies and magazines
(Morry \& Staska, 2001), or when wearing sexualizing clothing (B. L.
Fredrickson, Roberts, Noll, Quinn, \& Twenge, 1998). This momentary
self-objectification is referred to as \emph{state} self-objectification
(Bernard, Legrand, \& Klein, 2018; Calogero et al., 2011; Moradi \&
Huang, 2008). Being objectified by another person may interact TSO to
effect experiences of feeling like a body, or state self-objectification
(SSO; Garcia et al., 2016).

\subsection{Women Objectifying Women}\label{women-objectifying-women}

Harsey and Zurbriggen (2020) found that self-objectification was related
to the objectification of women to a similar degree for men and women.
Additionally, Strelan and Hargreaves (2005) found that the more women
self-objectify, the more they objectify other women. Could the process
work in the other direction as well, that is, could objectification by a
woman produce self-objectification, in a similar manner to mixed-gender
interactions?

Past research has focused on men's objectifying behaviors, with little
attention paid to the role that women may also play in objectifying
other women. Hill and Fischer (2008) assessed women's experiences of
objectification from men independently from women's experiences of
objectification from other women. They found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well. This process, however, was not comfirmed
to occur during an actual interpersonal interaction. Loughnan et al.
(2015a) found that women objectify other women to a greater extent than
they objectify men. Women have also been found to objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but very little is known about interpersonal sexual
objectification among women and what the proximal effects of this
objectification might be. We do know that objectification has more
adverse consequences for women than men (Gervais et al., 2011; Moradi \&
Huang, 2008; Saguy, Quinn, F Dovidio, \& Pratto, 2010), however, if
women objectify other women, perhaps even to a stronger degree than they
objectify men, this woman to woman objectification might have
detrimental outcomes.

\subsection{Interpersonal
Objectification}\label{interpersonal-objectification}

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Moreover, perhaps the most adverse effect is that being
objectified socializes girls and women to treat themselves as objects to
be looked at and evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr,
1972; B. L. Fredrickson et al., 1998). Gervais et al. (2020) recently
reviewed the research on interpersonal objectification and organized our
current theoretical understanding of this process in the Social
Interaction Model of Objectification (SIMO). What is clear from Gervais
et al. (2020)'s review is that we know little about women-on-women
interpersonal objectification. Could there also be negative consequences
of this type of interpersonal objectification?

Interpersonal objectification can also occur within the context of
romantic relationships, in addition to interactions with strangers.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, suggesting that women
can sometimes have positive outcomes from appearance valuation within
the context of an established romantic relationship. There is also
evidence that they experience less SSO after appearance comments within
romantic relationships (Meltzer, 2020), but less is known about first
dates. An important caveat to this literature is that it is only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) (Terán, Jiao, \& Aubrey,
2020). Both romantic relationships and stranger encounters have been
studied in the context of objectification, but neither of these
literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? On the one hand, the ample
research demonstrating that the male gaze has a particularly detrimental
effect would suggest no (Gay \& Castano, 2010; Gervais et al., 2011;
Saguy et al., 2010, Calogero (2004); Yilmaz \& Bozo, 2019), but on the
other hand, women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013). When they do it, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

\subsection{Interpersonal Objectification and
Authenticity}\label{interpersonal-objectification-and-authenticity}

There is some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán et al., 2020). This
link has been justified by the literature on stigmatized-stigmatizer
interactions (M. R. Hebl \& Dovidio, 2005; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), viewing the experience of being
objectified in an interaction as an identity threat situation
(Miles-McLean et al., 2015; Nadal \& Haynes, 2012). Further, empirical
evidence reveals that objectification manifests through inauthenticity
in romantic relationships (Brunell et al., 2010), adverse attitudes in
regard to career aspirations, and a decrease in concentration and
impairment in female cognitive performance (Kahalon, Shnabel, \& Becker,
2018; D. M. Quinn, Chaudoir, \& Kallen, 2011). When a woman is
objectified by a man, and subsequently experiences self-objectification,
the intergroup nature of this encounter might trigger identity threat.
Thus, perhaps there are fewer negative consequences when a woman is
objectified by another woman, specifically, woman-woman interpersonal
objectification processes might diverge from mix-gender interpersonal
objectification processes precisely because they are not intergroup
interactions (at least with respect to gender identity). A woman
objectified by another woman may not be having the same negative
consequences that cascade from situations that trigger group-based
identity threat (Deaux \& Major, 1987; Dovidio, Hebl, Richeson, \&
Shelton, 2006; M. R. Hebl \& Dovidio, 2005), but feeling like a body,
rather than a full human, in any interaction (intergroup or ingroup) may
be enough to reduce women's feelings of authenticity and social
competence, regardless of the gender of the objectifyer (Terán et al.,
2020; Tolman, Impett, Tracy, \& Michael, 2006).

\subsection{The Current Study}\label{the-current-study}

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected that TSO
would moderate this relationship, amplifying the positive assocaition
between being objectified and SSO for women higher in TSO. We
hypothesized that SSO would, in turn, lead to feelings of inauthenticity
(although there is ample support against this prediction from the
identity threat within intergroup interactions literature; Rollero,
2016). Further, we hypothesized that these feelings of inauthencity
would then be associated with reduced feelings of agency in romantic
relationships, reduced career aspiration, and reduced cognitive
performance. In summary, we expected to find a positive relationship
between other-objectification by one's partner and state
self-objectification. We also expected to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

These hypotheses are considered exploratory, given the lack of prior
research theorizing about these linkages.

Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would predict the TSO to other-objectification
link, further theorizing that this relationships is mediated by the
woman's own state self-objectification (SSO). This limited, but extant,
set of literature on woman-woman objectification points to a possible
alternative model from the model hypothesized in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

\section{Methods}\label{methods}

\subsection{Procedure}\label{procedure}

Except for the instructions given to participants, the procedure used
was identical to that in Garcia et al. (2016). In brief, that
methodology is that each participant arrived at the laboratory and were
then led into separate cubicles to prevent any communication between the
participants before the interaction. In addition, each participant was
screened for prior acquaintance to confirm that they had not met prior
to the study. They were asked to sign the consent form to participate,
and the study was described as follows: \enquote{This is a study looking
at how students form different types of relationships at college.} A
prompt on the computer screen told the participants that they were
assigned to the \enquote{College Relationships} condition and gave the
following instructions:

\begin{quote}
There are many types of relationships people form in college. During the
interaction, please think about your partner's potential as a romantic
partner. \textbf{Even if they are not the gender you are attracted to},
you can still judge their potential as a romantic partner. After the
interaction you will be asked to evaluate how dateable your partner is.
In other words, we would like to know if you think someone would date
your interaction partner. Also, your interaction partner will be
evaluating you in the same manner.
\end{quote}

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationship, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners (i.e., their sexual attractiveness; Puvia \& Vaes,
2013)---indeed, women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspect this this habitual
thought pattern will translate to their thoughts about other women.

After recieving the \enquote{College Relationships} prompt, the two
participants were then brought into a larger interaction room where they
sat on stools prearranged to be approximately 1 meter apart. The
experimenter instructed the participants to \enquote{get to know each
other} for 10 minutes and then left the room. After 10 minutes, the
experimenter came back into the room and stopped the interaction. The
participants then went back to their individual cubicles and completed a
set of post-interaction measures. Participants were then thanked for
their participation and debriefed. More detail on this methodology can
be found in Garcia et al. (2016).

\subsection{Participants}\label{participants}

Data from two demographically equivalent higher education institutions
in the Northeast United States were combined to create the final
analysis sample (\emph{N =} 64) used in this study. In the measures
section that follows we refer to them as Sample 1 and Sample 2.
Thirty-two previously unacquainted self-identifying female dyads (64
total participants) from institutions in the Northeast of the United
States participated in this study. Sample 1 (\emph{N =} 24) is from a
co-ed liberal arts college and Sample 2 (\emph{N =} 40) is from a
women's liberal arts college. More specifically, twelve of the pairs
were students at a co-ed liberal arts college, while the remaining
twenty pairs attended a women's liberal arts college. {[}Footnote:
Initially, data was collected from same-sex and mixed-sex dyads that
comprised of male and female gendered individuals. Sample 1 originally
consisted of twenty-two pairs, twelve men and thirty-two women.
Twenty-three pairs made up of forty-three women and one man, as well as
two participants who did not identify with either gender category,
formed Sample 2. For consistency, we limited participant data to same
sex female pairs at the two colleges.{]} All results presented below are
from models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were cross-racial minority group pairs. 64.06\%
of the sample identified as heterosexual, and 25\% identified as gay,
lesbian or bisexual.

\subsection{Post interaction Measures}\label{post-interaction-measures}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:corrtable}Correlations among study variables.}
\begin{tabular}{llllllll}
\toprule
 & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5}\\
\midrule
Actor's trait self objectification (TSO) & -0.35 & 2.64 &  &  &  &  & \\
Actor's authenticity of interaction & 5.23 & 1.02 & -.02 &  &  &  & \\
Actor's objectification of partner & -1.58 & 1.21 & .20 & -.07 &  &  & \\
Actor's state self-objectification & 1.92 & 1.13 & .13 & -.10 & -.09 &  & \\
Actor's future relationship agency & 4.69 & 0.96 & .04 & .23+ & .09 & -.09 & \\
Actor's cognitive performance & 5.03 & 2.29 & .08 & .11 & .11 & .02 & .07\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:descriptives}Descriptive Statistics for Study Variables}
\begin{tabular}{lll}
\toprule
 & \multicolumn{1}{c}{M} & \multicolumn{1}{c}{SD}\\
\midrule
Actor's trait self objectification (TSO) & -0.35 & 2.64\\
Actor's authenticity of interaction & 5.23 & 1.02\\
Actor's objectification of partner & -1.58 & 1.21\\
Actor's state self-objectification & 1.92 & 1.13\\
Actor's future relationship agency & 4.69 & 0.96\\
Actor's cognitive performance & 5.03 & 2.29\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

The following measures were collected in the order they are presented
following the interaction. Correlations appear in
Table~\ref{tab:corrtable}, and descriptive statistics appear in
Table~\ref{tab:descriptives}.

\subsubsection{Cognitive Performance}\label{cognitive-performance}

Trigrams from the Remote Associates Task (McFarlin \& Blascovich, 1984)
were utilized to assess cognitive performance after the interaction. Ten
items were selected and presented to participants. For example, the
correct answer for the trigram \enquote{Quack: Pond: Waddle} would be
\enquote{Duck}. Participants are limited to 30 seconds. For every
correct answer, 1 point is given. The mean score was 5.03 (SD = 2.29).
Cognitive performance was measured first in order to measure potential
immediate detriments to performance (Garcia et al., 2016).

\subsubsection{State
Other-Objectification}\label{state-other-objectification}

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner (Garcia et al., 2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were rated on a
scale from 1 (not at all) to 7 (constantly). Objectification was
measured by taking the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

As can be seen in Table~\ref{tab:descriptives}, the mean
other-objectification of women by women was \emph{M =} -1.58 (\emph{SD
=} 1.21). This corresponds to women objectifying other women to a
\emph{greater} extent than women's objectification of men reported in
Garcia et al. (2016) (\emph{M =} -1.68, \emph{SD =} 1.52). Further, in
the current sample the difference in other-objectification between
heterosexual (\emph{M =} -1.77, \emph{SD =} 1.14) and non-heterosexual
women (\emph{M =} -1.39, \emph{SD =} 1.35) was not statistically
significant, \emph{t}(25.89) = 1.02, \emph{p =} 0.32.

\subsubsection{Interaction Authenticity}\label{interaction-authenticity}

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much alike (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{`Do you think your partner
was authentic during your interaction?}' and \enquote{`Were you
authentic during your interaction?}' These questions were ranked on a
scale from 1 (not at all) to 7 (very much). These were combined to form
the authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) =
0.91 for Sample 2).

\subsubsection{State
Self-Objectification}\label{state-self-objectification}

To assess state self-objectification, we used an average of two items
from Saguy et al. (2010) that were also used in Garcia et al. (2016).
Participants were asked to rank how much they agreed with the following
statements: \enquote{During the interaction I felt more like a body than
a full self} and \enquote{I felt more like a body than as a real person
in the interaction}. Originally, Saguy et al. (2010) used 3 items, but
in both samples the reliability of the scale was higher once the third
item was removed, so we chose to only use the first two for our measure
of SSO, leaving us with a reliable scale (\(\alpha\) = 0.84 for Sample
1, and \(\alpha\) = 0.85 for Sample 2.)

\subsubsection{Relationship Agency}\label{relationship-agency}

A scale was used from Garcia et al. (2016) to assess how much agency an
individual believes they would possess in future romantic relationships.
Participants were asked how likely it was that they would do the
following: \enquote{`ask someone out on a date,}' \enquote{`open the
door for your date,}' \enquote{`pay for a date,}' \enquote{`ask your
boyfriend/girlfriend to marry you,}' \enquote{`initiate sex with your
girlfriend/boyfriend,}' \enquote{`initiate condom use during sex,}'
\enquote{`surprise your boyfriend/ girlfriend with a gift,}' and
\enquote{`ask your girlfriend/boyfriend to move with you to a new
place.}' Responses were measured on a scale ranging from 1 (not at all
likely) to 7 (extremely likely). The scale originally had 9 items, but
the 9th item had low correlations with the remaining items, ranging from
.02 to .30 for the first sample, and .04 to .30 for the second sample.
The item was intended to be reverse coded, but correlations were still
low enough to make the scale unreliable. Therefore, the ninth item was
removed. As a result, the scale had moderately high reliability for both
samples (\(\alpha\) = 0.72 for Sample 1, (\(\alpha\) = 0.74 for Sample
2).

\subsubsection{Career Aspirations}\label{career-aspirations}

To conceptualize participants' career aspirations after the interaction,
we used the 10-item adaptation of P. Gray and M. OBrien (2007)'s Career
Aspiration Scale employed in Garcia et al. (2016), which asked
participants to consider how true 10 statements were in regard to their
future careers on a scale from 0 (not at all true of me) to 4 (very true
of me). Items include \enquote{I hope to become a leader in my career
field} and \enquote{I hope to move up through any organization or
business I work in.} Items were fairly reliable (\(\alpha\) = 0.73 for
Sample 1, \(\alpha\) = 0.80 for Sample 2).

\subsubsection{Trait
Self-Objectification}\label{trait-self-objectification}

Trait self-objectification (TSO) was assessed using the
Self-Objectification Questionnaire (B. L. Fredrickson et al., 1998; M.
Noll \& L. Fredrickson, 1998), which evaluates the extent to which
individuals view their bodies in observable versus non-observable ways.
The questionnaire asked participants to rank order both appearance and
functional aspects of their bodies, from 1 (least important) to 10 (most
important), with respect to physical self-concepts. Of the ten body
attributes, five of the items were appearance-based (weight, sex appeal,
physical attractiveness, firm/sculpted muscles and body measurements),
and five of the items were competence-based (strength, physical
coordination, energy level, health and physical fitness). Difference
scores were computed by subtracting the sum of the 5 functional
aspects/competence attributes (e.g., health, strength) from the sum of
the 5 physical self-concepts/appearance attributes (e.g., physical
attractiveness, weight), and all measures were multiplied by -1, as was
done in Garcia et al. (2016), so that positive scores indicated greater
TSO.

\section{Results}\label{results}

\subsection{Data analysis}\label{data-analysis}

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{tidyverse} (Version 1.2.1; Wickham, 2017), \emph{psych} (Version
1.8.4; Revelle, 2017), \emph{nlme} (Version 3.1.137; Pinheiro, Bates,
DebRoy, Sarkar, \& R Core Team, 2017) for our analyses. Further, we used
the R-packages \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018),
\emph{apaTables} (Version 2.0.5; Stanley, 2018), with \emph{knitr}
(Version 1.25; Xie, 2015) to create a fully reproducible manuscript in
\emph{rmarkdown} (Version 1.15; Xie, Allaire, \& Grolemund, 2018).

\subsection{Analysis Strategy}\label{analysis-strategy}

\begin{figure}
\includegraphics[width=400px]{APIM_figure} \caption{Basic actor-partner interdependence model (APIM) depiction.}\label{fig:apim}
\end{figure}

\begin{figure}
\includegraphics[width=400px]{theor_figure} \caption{Path Analysis Model from Garcia et al. (2016) study with distinguishable dyads.}\label{fig:theorfigure}
\end{figure}

This study sought to detect whether partners' objectification of one
another affected state self-objectification (SSO). Specifically, we were
interested in testing the relationship between state-other
objectification and SSO, and how SSO in turn, affects feelings of
inauthenticity during the interaction. In addition, we also tested if
the effect of other-objectification in an interaction on SSO is only
present for those women who are high in trait self-objectification.
Further, we investigated the relationships between experiencing
interaction inauthenticity and relationship agency, career aspirations,
and cognitive performance.

While past studies investigating objectification in interactions used
dyadic path analysis (Garcia et al., 2016), the current study used
multilevel modeling proceedures. Dyadic analyses for distinguishable
dyads (e.g., mixed-gender interacting pairs) is more natural in
Structural Equation Modeling (SEM) than it is for indistinguishable
dyads (e.g., same-gender interacting pairs), as we have here (Garcia,
Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One reason for
this asymmetry is that, due to the arbitrary distinctions made between
\enquote{partner 1} and \enquote{partner 2} in indistinguishable dyads,
many estimates need to be fixed to be equal (i.e., paths, variances,
covariances, endogenous intercepts, and exogenous means), but these
equality constraints should not then be considered in the degrees of
freedom calculations for fit estimations (Olsen \& Kenny, 2006).
Further, Olsen and Kenny (2006) details how a new independence model and
the corresponding fit measure should be re-calculated for
indistinguishable dyads models. The current study uses dyadic multilevel
modeling (MLM) to test all relationships and mediation patterns. The
online supplementary materials contains analysis using SEM. See
Ledermann and Kenny (2017) for a more complete discussion of the
considerations for using SEM versus MLM for dyadic analysis.

Testing our hypotheses in the current, same-gender, sample, involves
using the Actor-Partner Independence Model (APIM) approach for each
outcome variable (i.e., endogenous variable in
Figure~\ref{fig:theorfigure}). Thus, we ran five APIM's to test all the
hypothesized relationships. See Figure~\ref{fig:apim} for a basic APIM
model. The APIM includes effects due to one's own, as well as one's
partner's, predictor variables (\(X\)'s) on the one's own outcome
variable (\(Y\)). The current study deals with indistinguishable dyads,
meaning the designation of who is designated as \enquote{actor} and who
is designated as \enquote{partner} is arbitrary. These analyses are
considered exploratory, given the lack of prior research theorizing
about these linkages.

\subsection{Main Results}\label{main-results}

\begin{figure}
\includegraphics[width=400px]{SEMfigure} \caption{Path Analysis Model with Estimates}\label{fig:semfigure}
\end{figure}

All model estimates and p-values are found in Table~\ref{tab:mlm} and
the relationships with estimates included are depicted in
Figure~\ref{fig:semfigure}.

The most important finding from Garcia et al. (2016) was the significant
partner effect of other objectification and SSO (specifically men's
objectification of women and women's SSO). As expected, the partner
effect of other-objectification on SSO in the current all-women sample
was statistically significant, \emph{b} = 0.29, \emph{SE} = 0.12,
\emph{p} = .019. One's own other objectification had no effect on SSO,
\emph{b} = -0.16, \emph{SE} = 0.12, \emph{p} = .210. Inconsistent with
past findings however, there was no statistically significant
interaction of partner's other objectification and the person's trait
self-objectification on SSO, \emph{b} = 0.03, \emph{SE} = 0.27, \emph{p}
= .910. There was also no significant main effect of trait
self-objectification on SSO \emph{b} = 0.07, \emph{SE} = 0.05, \emph{p}
= .183.

There was no significant effect of SSO on interaction authenticity,
although the estimate of this effect was in the hypothesized negative
direction, \emph{b} = -0.09, \emph{SE} = 0.12, \emph{p} = .431. Because
authenticity was a composite score of 9 items, two of which were
interaction specific authenticity items, we also estimated the pairwise
correlations between SSO and all these items individually. They were all
small, ranging from only -0.01 to -0.14. Although we hypothesized that
SSO would mediate the relationship between partner's other
objectification and interaction authenticity, after finding no
relationship between SSO and authenticity, we also tested if the
partner's other objectification had a direct effect on authenticity, but
this effect was not significant, \emph{b} = 0.06, \emph{SE} = 0.12,
\emph{p} = .608 (nor was the total effect of partner's other
objectification on authenticity, \emph{b} = 0.03, \emph{SE} = 0.11,
\emph{p} = .787).

Lastly, although there was no evidence that SSO was related to
interaction authenticity in the current sample, we tested if interaction
authenticity (composite of nine items) had effects on cognitive
performance, career aspirations, and relationship agency, as it did in
Garcia et al. (2016). We again used MLM and thus, these effects were
tested in three separate models. There was no significant effect of
interaction authenticity on cognitive performance, \emph{b} = 0.32,
\emph{SE} = 0.28, \emph{p} = .258, but authenticity was significantly
positively related to both career aspirations, \emph{b} = 0.18,
\emph{SE} = 0.07, \emph{p} = .010, and relationship agency, \emph{b} =
0.23, \emph{SE} = 0.12, \emph{p} = .049. There was no direct effect of
SSO on cognitive performance, \emph{b} = 0.04, \emph{SE} = 0.25,
\emph{p} = .872, and no direct effect of partner's other objectification
on cognitive performance, \emph{b} = 0.01, \emph{SE} = 0.27, \emph{p} =
.962. There was no direct effect of SSO on career aspirations, \emph{b}
= 0.03, \emph{SE} = 0.06, \emph{p} = .657, and no direct effect of
partner's other objectification on career aspirations, \emph{b} = -0.06,
\emph{SE} = 0.07, \emph{p} = .378. There was no direct effect of SSO on
relationship agency, \emph{b} = -0.05, \emph{SE} = 0.11, \emph{p} =
.659, and no direct effect of partner's other objectification on
relationship agency, \emph{b} = -0.1, \emph{SE} = 0.11, \emph{p} = .365.

\section{Discussion}\label{discussion}

The current study tested whether the model of interpersonal
objectification and state self-objectification (SSO) used in Garcia et
al. (2016) replicates in a same-gender dyadic sample of women only.
Although past research has found that women do objectify other women
(Harsey \& Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes,
2013), this is the first study to test if \emph{interpersonal}
other-objectification by women is related to state self-objectification
in their woman-identified interaction partners. As hypothesized, the
current study did find a significant relationship between a woman's
interaction partner's report of having objectified her and her own
post-interaction feelings of self-objectification. That is, there was a
significant partner effect of other-objectification on SSO. This effect
extends the equivalent relationship found in mixed-gender interactions
to the context of same-gender interactions between women. Thus, evidence
suggests that it is not only the male gaze, and male
other-objectification that is related to women's state
self-objectification, but being objectified by another woman can also
result in SSO, at least in the context of a scenario where they know
they are being evaluated as a potential dating partner. {[}Asha:Or
\ldots{} thus evidence provides support for the claim that women being
objectified by another woman result in sso in ways that are not uniquely
dissimilar from that of xyz{]}

As is the case in all correlational studies, we cannot be sure about the
causal direction. It could be that women's SSO causes them to be
objectified by their interaction partner. This interpretation is
theoretically interesting given that, in both the Garcia et al. (2016)
study and in the current study, it is \emph{women's} SSO that relates to
being objectified by one's partner. However, Objectification Theory
(Roberts \& Fredrickson, 1997), as well as past experimental studies
{[}Saguy et al. (2010); CITE{]}, suggest that the causal flow is from
other-objectification to SSO. Although, the current study found evidence
for the partner effect of other-objectification on SSO, there was a lack
of evidence for the connections between SSO and downstream negative
consequences.

Where the results of the current study diverge most notably from the
results of studies testing interpersonal objectification among
mixed-gender dyads is the lack of evidence for relationships between SSO
and interaction inauthenticity. In addition, we found no evidence of
\emph{direct} or \emph{indirect} relationships between SSO and cognitive
functioning, relationship agency, and career aspirations. Nor was there
evidence of \emph{direct} or \emph{indirect} relationships between
other-objectification and any of these outcomes. This is somewhat
suprising given the extant evidence linking SSO and cognitive
functioning (B. L. Fredrickson et al., 1998; and see Moradi \& Huang,
2008 for a review; Quinn, Chaudoir, \& Kallen, 2011) and the small, but
extant research on interpersonal other-ojectification and cognitive
functioning (Garcia et al., 2016; Logel et al., 2009). This lack of
evidence could potentially signal diverging processes between women's
experiences with interpersonal objectification from men and
interpersonal objectification from women. There is quite a bit of
evidence suggesting that the male gaze is particularly detrimental
(Calogero, 2004; Gervais, Holland, \& Dodd, 2013; Gervais et al., 2011;
Roberts \& Fredrickson, 1997), and perhaps the self-objectification
experienced within an interaction with a women is not as harmful as the
self-objectification experienced within an interaction with a man. As a
strong note of caution, it is best not to interpret a null result as
evidence of no relationship.

The lack of evidence for a relationship between SSO and interaction
authenticity is surprising and again, should not be interpreted as
evidence of no relationship. It should be noted that the estimate of
this relationship was very small, but in the negative direction, as
anticipated. If it is the case that there is a smaller (i.e., weaker)
connection between women's feelings of SSO and inauthenticity in
interactions with other women than in interactions with men, models of
interpersonal objectification, like the SIMO (Gervais et al., 2020),
could be extended by including inauthenticity as a potential mediating
factor. Further, gender of the objectifyer/interaction partner could
also be added to the SIMO model to help understand the circumstances
that require other-objectification and SSO to have negative consequences
for women.

Although we did not find a connection between SSO and authenticity, we
did find significant positive relationships between authenticity and
relationship agency, and authenticity and career aspirations. The
relationship between authenticity and cognitive functioning was also
estimated as positive, but was not statistically signficant. Again, due
to the lack of connection between SSO and authenticity, we found no
evidence of \emph{indirect} relationships between SSO and these outcome
variables. This evidence of a relationship between authenticity and the
relational outcome variables (i.e., relationship agency and career
aspirations) provides evidence that corroborates past findings that felt
authenticity in interactions is important for healthy relationship
functioning (Garcia et al., 2016; Terán et al., 2020). Just as
authenticity has been found to be important in intergroup interactions
(Brunell et al., 2010; Garcia et al., 2016; Tolman et al., 2006), we
again find more evidence here that disruptions in feelings of
authenticity can negatively impact relationships beyond the current
partner. Although the current study did not find a connection between
authenticity and SSO, this seems theoretically to be a natural
connection, and future work must explore when and how SSO leads to
inauthenticity in interpersonal objectification situations.

\subsection{Limitations and Future
Directions}\label{limitations-and-future-directions}

\subsubsection{Sample Characteristics}\label{sample-characteristics}

In addition to being small and combined across two institutions, the
sample of the current study was comprised of Western women, being that
sexual objectification is most prevalent in this culture (Loughnan et
al., 2015b), and research on objectification conducted outside of
Western or Westernized countries is scarce (Moradi \& Huang, 2008),
although more current work examines objectification from a
cross-cultural framework (Loughnan et al., 2015b, Wollast et al.
(2020)). Since \enquote{bodies exist within social and cultural
contexts, and hence are also constructed through sociocultural practices
and discourses} (Roberts \& Fredrickson, 1997, p. 174), it is important
to consider how diverse social identities within unique cultural
contexts may inform sexual objectification phenomenon to test the
cross-cultural applicability of theoretical frameworks (Loughnan et al.,
2015b). Further, sexualizing experiences and self-objectification are
thought to begin a very young age, and thus, researchers have only
recently begun to examine such experiences among children (Bury,
Tiggemann, \& Slater, 2016; e.g., Holland \& Haslam, 2016; Jongenelis,
Byrne, \& Pettigrew, 2014). Considering the fact that the average mean
age of the investigated participants of this current study was 18.85
years, research among younger and older individuals is needed,
especially because self-objectification may change over time (Roberts \&
Fredrickson, 1997). It may be valuable to question the extent to which
children, adolescents, or emerging adults of different races or
ethnicities are exposed to varied amounts of sexualizing content.
{[}Kat: Or a longitudinal study is needed.{]}

\subsubsection{Sexual Objectification and the Objectifyer's
Gender}\label{sexual-objectification-and-the-objectifyers-gender}

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential dating/romantic partner. We
think that hetereosexual women are able to do this with other women as
their target---and there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---they may be even
more apt to activate social comparison processes (Festinger, 1954) than
women who are sexually attracted to other women (non-heterosexual
women). This differential psychological process between women with
differing sexualities might have served to dampen our ability to detect
relationships, adding variability. {[}Kat: The only other thing I would
add re: sexuality limitations are that we didn't collect/use sexual
orientation, so we couldn't analyze results by that at all/our samples
might have differed there.{]}

Previous research has found that when compared to heterosexual women,
lesbian women report less concern with physical appearance (Siever,
1994; Strong, Williamson, Netemeyer, \& Geer, 2000), and less
self-objectification (Brownlow, 1998; Noffsinger-Frazier, 2004). However
none of these studies examined the relationship between
self-objectification and experiences of sexual objectification. Thus, it
is unclear whether lesbians indeed experience similar levels of cultural
sexual objectification but internalize them less than heterosexual women
do. Consistent with previous research, Hill and Fischer (2008)
determined that lesbians existed less physical appearance concerns
compared to heterosexual women, however there was no evidence found that
lesbian women self-objectify less than heterosexual women and they did
not find that sexual orientation moderates the relationship between
sexual objectification and self-objectification. This contradicts older
theoretical literature that suggests that lesbians internalize cultural
sexual objectification less than do heterosexual women (Brown, 1987;
LUM, 1994; Pitman, 1999; Siever, 1994).

However, since both men and women are socialized in a culture that
sexually objectifies women, both men and women may come to internalize
this socialization and sexually objectify women. Indeed, recent research
has found that the more women self-objectify, the more they objectify
other women (Loughnan et al., 2015a; Puvia \& Vaes, 2013), although not
to the degree exhibited by men; that is, men were found to objectify
women significantly more than women objectify other women (Strelan \&
Hargreaves, 2005).

Objectification among platonic friends: Perhaps women could be
amplifying each others' state self-objectification \emph{before} the
theoretical encounters with men organized in the SIMO occur. In this
way, women's rituals of getting ready together might place even women
with low TSO in a higher SSO group.

\subsubsection{Sexism and Sexual
Discrimination}\label{sexism-and-sexual-discrimination}

Sexism has recently been found to be related to the objectification of
women, but not men, for both men and women (Harsey \& Zurbriggen,
2020).{[}Kat: Is the relationship between sexism and objectification
really this recent of a finding? Or was it just replicated recently?{]}
Empirical evidence illustrates how women continue to be objects of
interpersonal discrimination and experience daily sexist hassles (Swim,
Hyers, Cohen, \& Ferguson, 2001). One form of interpersonal
discrimination women face is the process by which their whole being is
viewed as a collection of sexualized body parts valued predominantly for
commodification, a phenomena termed sexual objectification (Bartky,
1990). Sexual objectification occurs with both \enquote{endless variety
and monotonous similarity,} and is thus mediated by unique combinations
of race, ethnicity, sexuality, age, and class (Fredrickson, Hendler,
Nilsen, O'Barr, \& Roberts, 2011; Rubin, 1975, cited in Fraser and
Nicholson (1989), p.~28). Amid such heterogeneity though,
\enquote{having a reproductively mature female body} proposed by Roberts
and Fredrickson (1997) is likely to create a shared vulnerability to
sexual objectification and a variety of shared negative experiences as a
result.

\subsubsection{Measurement}\label{measurement}

Future experiments or longitudinal studies should explore the external
validity of the notions of self-objectification and how the
operationalization of self-objectification may be improved.

\subsubsection{Alternative Model}\label{alternative-model}

Puvia and Vaes (2013) alternative model. TSO -\textgreater{} SOO,
mediated by SSO. \enquote{Found that women's motivation to look
attractive (for men) and internalize sociocultural beauty standards were
positively linked with the dehumanization of sexually objectified female
targets.} Tis relationship was mediated by their own level of
self-objectification.''

\subsection{Clinical Implications}\label{clinical-implications}

The results from the current analysis highlight how subtle forms of
sexist discrimination operate, and provide valuable insight into
prevention and intervention efforts in both clinical and educational
contexts. These results are quite useful for promoting mental health,
the creation and maintence of early action programs for girls and young
women, and for scholars and practitioners to work intentially to provide
the tools necessary to circumvent or mitigate negative effects on
self-objectification to combat such experiences.

\section{Conclusion}\label{conclusion}

The results of this study demonstrate the complex and ambivalent nature
of sexual objectification and additionally highlight the psychological
and social consequences of such objectification processes on women's
social relationships and well-being. We did not find a significant
effect between actor SSO and felt authenticity in the interaction, which
suggests that there is not sufficient evidence to support the claim that
partner objectification is the cause for the diverse range of negative
effects related to interaction inauthenticity.

\section{References}\label{references}

\newpage

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-argyle1969}{}
Argyle, M., \& Williams, M. (1969). Observer or observed? A reversible
perspective in person perception. \emph{Sociometry}, 396--412.

\hypertarget{ref-R-papaja}{}
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Bartky}{}
Bartky, S. L. (1990). Femininity and domination studies in the
phenomenology of oppression.

\hypertarget{ref-berger1972}{}
Berger, J., Cohen, B. P., \& Zelditch Jr, M. (1972). Status
characteristics and social interaction. \emph{American Sociological
Review}, 241--255.

\hypertarget{ref-bernard2018bodies}{}
Bernard, P., Legrand, S., \& Klein, O. (2018). From bodies to blame:
Exposure to sexually objectifying media increases tolerance toward
sexual harassment. \emph{Psychology of Popular Media Culture},
\emph{7}(2), 99.

\hypertarget{ref-briton1995}{}
Briton, N. J., \& Hall, J. A. (1995). Beliefs about female and male
nonverbal communication. \emph{Sex Roles}, \emph{32}(1-2), 79--90.

\hypertarget{ref-brown1987lesbians}{}
Brown, L. S. (1987). Lesbians, weight, and eating: New analyses and
perspectives. \emph{Lesbian Psychologies: Explorations and Challenges},
294--309.

\hypertarget{ref-brownlow1998relationship}{}
Brownlow, B. S. (1998). \emph{The relationship between objectification,
body image disturbances, and disordered eating: Investigating race,
socioeconomic status, acculturation and self-objectification as
mediators.} (PhD thesis). ProQuest Information \& Learning.

\hypertarget{ref-brunelletal2010}{}
Brunell, A. B., Kernis, M. H., Goldman, B. M., Heppner, W., Davis, P.,
Cascio, E. V., \& Webster, G. D. (2010). Dispositional authenticity and
romantic relationship functioning. \emph{Personality and Individual
Differences}, \emph{48}(8), 900--905.
doi:\href{https://doi.org/10.1016/j.paid.2010.02.018}{10.1016/j.paid.2010.02.018}

\hypertarget{ref-bury2016}{}
Bury, B., Tiggemann, M., \& Slater, A. (2016). Disclaimer labels on
fashion magazine advertisements: Impact on visual attention and
relationship with body dissatisfaction. \emph{Body Image}, \emph{16},
1--9.

\hypertarget{ref-calogero2004test}{}
Calogero, R. M. (2004). A test of objectification theory: The effect of
the male gaze on appearance concerns in college women. \emph{Psychology
of Women Quarterly}, \emph{28}(1), 16--21.

\hypertarget{ref-calogero2011}{}
Calogero, R. M., Tantleff-Dunn, S. E., \& Thompson, J. (2011).
\emph{Self-objectification in women: Causes, consequences, and
counteractions.} American Psychological Association.

\hypertarget{ref-deaux1987putting}{}
Deaux, K., \& Major, B. (1987). Putting gender into context: An
interactive model of gender-related behavior. \emph{Psychological
Review}, \emph{94}(3), 369.

\hypertarget{ref-dovidio2006nonverbal}{}
Dovidio, J. F., Hebl, M., Richeson, J. A., \& Shelton, J. N. (2006).
Nonverbal communication, race, and intergroup interaction.

\hypertarget{ref-festinger1954theory}{}
Festinger, L. (1954). A theory of social comparison processes.
\emph{Human Relations}, \emph{7}(2), 117--140.

\hypertarget{ref-frasernicholson1989}{}
Fraser, N., \& Nicholson, L. (1989). Social criticism without
philosophy: An encounter between feminism and postmodernism.
\emph{Social Text}, (21), 83.
doi:\href{https://doi.org/10.2307/827810}{10.2307/827810}

\hypertarget{ref-fredricksonetal2011}{}
Fredrickson, B. L., Hendler, L. M., Nilsen, S., O'Barr, J. F., \&
Roberts, T.-A. (2011). Bringing back the body. \emph{Psychology of Women
Quarterly}, \emph{35}(4), 689--696.
doi:\href{https://doi.org/10.1177/0361684311426690}{10.1177/0361684311426690}

\hypertarget{ref-fredrickson1998swimsuit}{}
Fredrickson, B. L., Roberts, T.-A., Noll, S. M., Quinn, D. M., \&
Twenge, J. M. (1998). That swimsuit becomes you: Sex differences in
self-objectification, restrained eating, and math performance.
\emph{Journal of Personality and Social Psychology}, \emph{75}(1), 269.

\hypertarget{ref-garcia2016objectification}{}
Garcia, R. L., Earnshaw, V. A., \& Quinn, D. M. (2016). Objectification
in action: Self-and other-objectification in mixed-sex interpersonal
interactions. \emph{Psychology of Women Quarterly}, \emph{40}(2),
213--228.

\hypertarget{ref-garcia2015moderation}{}
Garcia, R. L., Kenny, D. A., \& Ledermann, T. (2015). Moderation in the
actor--partner interdependence model. \emph{Personal Relationships},
\emph{22}(1), 8--29.

\hypertarget{ref-gay2010my}{}
Gay, R. K., \& Castano, E. (2010). My body or my mind: The impact of
state and trait objectification on women's cognitive resources.
\emph{European Journal of Social Psychology}, \emph{40}(5), 695--703.

\hypertarget{ref-gervais2013my}{}
Gervais, S. J., Holland, A. M., \& Dodd, M. D. (2013). My eyes are up
here: The nature of the objectifying gaze toward women. \emph{Sex
Roles}, \emph{69}(11-12), 557--570.

\hypertarget{ref-gervais2020social}{}
Gervais, S. J., Sáez, G., Riemer, A. R., \& Klein, O. (2020). The social
interaction model of objectification: A process model of goal-based
objectifying exchanges between men and women. \emph{British Journal of
Social Psychology}, \emph{59}(1), 248--283.

\hypertarget{ref-gervais2011you}{}
Gervais, S. J., Vescio, T. K., \& Allen, J. (2011). When what you see is
what you get: The consequences of the objectifying gaze for women and
men. \emph{Psychology of Women Quarterly}, \emph{35}(1), 5--17.

\hypertarget{ref-harsey2020men}{}
Harsey, S. J., \& Zurbriggen, E. L. (2020). Men and women's
self-objectification, objectification of women, and sexist beliefs.
\emph{Self and Identity}, 1--8.

\hypertarget{ref-hebl2005promoting}{}
Hebl, M. R., \& Dovidio, J. F. (2005). Promoting the ``social'' in the
examination of social stigmas. \emph{Personality and Social Psychology
Review}, \emph{9}(2), 156--182.

\hypertarget{ref-hill2008examining}{}
Hill, M. S., \& Fischer, A. R. (2008). Examining objectification theory:
Lesbian and heterosexual women's experiences with sexual-and
self-objectification. \emph{The Counseling Psychologist}, \emph{36}(5),
745--776.

\hypertarget{ref-holland2016}{}
Holland, E., \& Haslam, N. (2016). Cute little things: The
objectification of prepubescent girls. \emph{Psychology of Women
Quarterly}, \emph{40}(1), 108--119.

\hypertarget{ref-jongenelis2014}{}
Jongenelis, M. I., Byrne, S. M., \& Pettigrew, S. (2014).
Self-objectification, body image disturbance, and eating disorder
symptoms in young australian children. \emph{Body Image}, \emph{11}(3),
290--302.

\hypertarget{ref-kahalon2018don}{}
Kahalon, R., Shnabel, N., \& Becker, J. C. (2018). ``Don't bother your
pretty little head'' appearance compliments lead to improved mood but
impaired cognitive performance. \emph{Psychology of Women Quarterly},
\emph{42}(2), 136--150.

\hypertarget{ref-ledermann2017analyzing}{}
Ledermann, T., \& Kenny, D. A. (2017). Analyzing dyadic data with
multilevel modeling versus structural equation modeling: A tale of two
methods. \emph{Journal of Family Psychology}, \emph{31}(4), 442.

\hypertarget{ref-logel2009interacting}{}
Logel, C., Walton, G. M., Spencer, S. J., Iserman, E. C., Hippel, W.
von, \& Bell, A. E. (2009). Interacting with sexist men triggers social
identity threat among female engineers. \emph{Journal of Personality and
Social Psychology}, \emph{96}(6), 1089.

\hypertarget{ref-loughnan2015exploring}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015a). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

\hypertarget{ref-loughnan2015}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015b). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

\hypertarget{ref-lum1994lesbians}{}
LUM, E. (1994). Lesbians and physical appearance which model applies?
\emph{Lesbian and Gay Psychology: Theory, Research, and Clinical
Applications}, \emph{1}, 84.

\hypertarget{ref-nollfredrickson1998}{}
M. Noll, S., \& L. Fredrickson, B. (1998). A mediational model linking
self-objectification, body shame, and disordered eating.
\emph{Psychology of Women Quarterly}, \emph{22}, 623--636.
doi:\href{https://doi.org/10.1111/j.1471-6402.1998.tb00181.x}{10.1111/j.1471-6402.1998.tb00181.x}

\hypertarget{ref-mcfarlin1984remote}{}
McFarlin, D. B., \& Blascovich, J. (1984). On the remote associates test
(rat) as an alternative to illusory performance feedback: A
methodological note. \emph{Basic and Applied Social Psychology},
\emph{5}(3), 223--229.

\hypertarget{ref-meltzer2020women}{}
Meltzer, A. L. (2020). Women can benefit from sexual and physical
valuation in the context of a romantic relationship. \emph{Personality
and Social Psychology Bulletin}, \emph{46}(2), 243--257.

\hypertarget{ref-miles2015stop}{}
Miles-McLean, H., Liss, M., Erchull, M. J., Robertson, C. M., Hagerman,
C., Gnoleba, M. A., \& Papp, L. J. (2015). ``Stop looking at me!''
interpersonal sexual objectification as a source of insidious trauma.
\emph{Psychology of Women Quarterly}, \emph{39}(3), 363--374.

\hypertarget{ref-moradi2008}{}
Moradi, B., \& Huang, Y.-P. (2008). Objectification theory and
psychology of women: A decade of advances and future directions.
\emph{Psychology of Women Quarterly}, \emph{32}(4), 377--398.

\hypertarget{ref-morry2001magazine}{}
Morry, M. M., \& Staska, S. L. (2001). Magazine exposure:
Internalization, self-objectification, eating attitudes, and body
satisfaction in male and female university students. \emph{Canadian
Journal of Behavioural Science/Revue Canadienne Des Sciences Du
Comportement}, \emph{33}(4), 269.

\hypertarget{ref-nadal2012effects}{}
Nadal, K. L., \& Haynes, K. (2012). The effects of sexism, gender
microaggressions, and other forms of discrimination on women's mental
health and development.

\hypertarget{ref-noffsinger2004objectification}{}
Noffsinger-Frazier, N. A. (2004). \emph{Objectification theory and
disordered eating: The impact of feminist identification,
internalization of sociocultural standards of appearance, and sexual
orientation}. The University of Memphis.

\hypertarget{ref-olsen2006structural}{}
Olsen, J. A., \& Kenny, D. A. (2006). Structural equation modeling with
interchangeable dyads. \emph{Psychological Methods}, \emph{11}(2), 127.

\hypertarget{ref-grayobrien2007}{}
P. Gray, M., \& M. OBrien, K. (2007). Advancing the assessment of
women's career choices: The career aspiration scale. \emph{Journal of
Career Assessment - J CAREER ASSESSMENT}, \emph{15}, 317--337.
doi:\href{https://doi.org/10.1177/1069072707301211}{10.1177/1069072707301211}

\hypertarget{ref-R-nlme}{}
Pinheiro, J., Bates, D., DebRoy, S., Sarkar, D., \& R Core Team. (2017).
\emph{nlme: Linear and nonlinear mixed effects models}. Retrieved from
\url{https://CRAN.R-project.org/package=nlme}

\hypertarget{ref-pitman1999body}{}
Pitman, G. E. (1999). Body image, compulsory heterosexuality, and
internalized homophobia. \emph{Journal of Lesbian Studies}, \emph{3}(4),
129--139.

\hypertarget{ref-puvia2013being}{}
Puvia, E., \& Vaes, J. (2013). Being a body: Women's appearance related
self-views and their dehumanization of sexually objectified female
targets. \emph{Sex Roles}, \emph{68}(7-8), 484--495.

\hypertarget{ref-quinnetal}{}
Quinn, D. M., Chaudoir, S. R., \& Kallen, R. W. (2011). Performance and
flow: A review and integration of self-objectification research.
\emph{Self-Objectification in Women: Causes, Consequences, and
Counteractions.}, 119--138.
doi:\href{https://doi.org/10.1037/12304-006}{10.1037/12304-006}

\hypertarget{ref-quinn2011performance}{}
Quinn, D. M., Chaudoir, S. R., \& Kallen, R. W. (2011). Performance and
flow: A review and integration of self-objectification research.

\hypertarget{ref-R-base}{}
R Core Team. (2017). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-R-psych}{}
Revelle, W. (2017). \emph{Psych: Procedures for psychological,
psychometric, and personality research}. Evanston, Illinois:
Northwestern University. Retrieved from
\url{https://CRAN.R-project.org/package=psych}

\hypertarget{ref-richeson2003prejudice}{}
Richeson, J. A., \& Shelton, J. N. (2003). When prejudice does not pay:
Effects of interracial contact on executive function.
\emph{Psychological Science}, \emph{14}(3), 287--290.

\hypertarget{ref-riemer2020self}{}
Riemer, A. R., Sáez, G., Brock, R., \& Gervais, S. J. (2020).
Self-fulfilling objectification in relationships: The effects of men's
objectifying expectations on women's self-objectification during
conflict in romantic relationships. \emph{Self and Identity}, 1--7.

\hypertarget{ref-robertsfredrickson}{}
Roberts, T.-A., \& Fredrickson, B. L. (1997). Objectification theory.
\emph{Encyclopedia of Social Psychology}.
doi:\href{https://doi.org/10.4135/9781412956253.n377}{10.4135/9781412956253.n377}

\hypertarget{ref-rollero2016bringing}{}
Rollero, C. (2016). Bringing objectification into social relationships
research: Is self-objectification harmful for authenticity?

\hypertarget{ref-rubin1975}{}
Rubin, J. (1975). What the ``good language learner'' can teach us.
\emph{TESOL Quarterly}, \emph{9}(1), 41.
doi:\href{https://doi.org/10.2307/3586011}{10.2307/3586011}

\hypertarget{ref-saguyetal2010}{}
Saguy, T., Quinn, D., F Dovidio, J., \& Pratto, F. (2010). Interacting
like a body: Objectification can lead women to narrow their presence in
social interactions. \emph{Psychological Science}, \emph{21}, 178--82.
doi:\href{https://doi.org/10.1177/0956797609357751}{10.1177/0956797609357751}

\hypertarget{ref-sanchez2008romance}{}
Sanchez, D. T., \& Broccoli, T. L. (2008). The romance of
self-objectification: Does priming romantic relationships induce states
of self-objectification among women? \emph{Sex Roles}, \emph{59}(7-8),
545--554.

\hypertarget{ref-shelton2005expecting}{}
Shelton, J. N., Richeson, J. A., \& Salvatore, J. (2005). Expecting to
be the target of prejudice: Implications for interethnic interactions.
\emph{Personality and Social Psychology Bulletin}, \emph{31}(9),
1189--1202.

\hypertarget{ref-siever1994sexual}{}
Siever, M. D. (1994). Sexual orientation and gender as factors in
socioculturally acquired vulnerability to body dissatisfaction and
eating disorders. \emph{Journal of Consulting and Clinical Psychology},
\emph{62}(2), 252.

\hypertarget{ref-R-apaTables}{}
Stanley, D. (2018). \emph{ApaTables: Create american psychological
association (apa) style tables}. Retrieved from
\url{https://CRAN.R-project.org/package=apaTables}

\hypertarget{ref-strelan2005women}{}
Strelan, P., \& Hargreaves, D. (2005). Women who objectify other women:
The vicious circle of objectification? \emph{Sex Roles},
\emph{52}(9-10), 707--712.

\hypertarget{ref-strelan2018birds}{}
Strelan, P., \& Pagoudis, S. (2018). Birds of a feather flock together:
The interpersonal process of objectification within intimate
heterosexual relationships. \emph{Sex Roles}, \emph{79}(1-2), 72--82.

\hypertarget{ref-strong2000eating}{}
Strong, S. M., Williamson, D. A., Netemeyer, R. G., \& Geer, J. H.
(2000). Eating disorder symptoms and concerns about body differ as a
function of gender and sexual orientation. \emph{Journal of Social and
Clinical Psychology}, \emph{19}(2), 240--255.

\hypertarget{ref-swimetal}{}
Swim, J. K., Hyers, L. L., Cohen, L. L., \& Ferguson, M. J. (2001).
Everyday sexism: Evidence for its incidence, nature, and psychological
impact from three daily diary studies. \emph{Journal of Social Issues},
\emph{57}(1), 31--53.
doi:\href{https://doi.org/10.1111/0022-4537.00200}{10.1111/0022-4537.00200}

\hypertarget{ref-teran2020relational}{}
Terán, L., Jiao, J., \& Aubrey, J. S. (2020). The relational burden of
objectification: Exploring how past experiences of interpersonal sexual
objectification are related to relationship competencies. \emph{Sex
Roles}, 1--16.

\hypertarget{ref-tolman2006looking}{}
Tolman, D. L., Impett, E. A., Tracy, A. J., \& Michael, A. (2006).
Looking good, sounding good: Femininity ideology and adolescent girls'
mental health. \emph{Psychology of Women Quarterly}, \emph{30}(1),
85--95.

\hypertarget{ref-R-tidyverse}{}
Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
'tidyverse'}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyverse}

\hypertarget{ref-wollast2020cultural}{}
Wollast, R., Riemer, A. R., Gervais, S. J., Grigoryan, L., Bernard, P.,
\& Klein, O. (2020). How cultural orientation and self-compassion shape
objectified body consciousness for women from america, belgium, russia,
and thailand. \emph{Self and Identity}, 1--21.

\hypertarget{ref-R-knitr}{}
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.).
Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://yihui.name/knitr/}

\hypertarget{ref-R-rmarkdown}{}
Xie, Y., Allaire, J., \& Grolemund, G. (2018). \emph{R markdown: The
definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved
from \url{https://bookdown.org/yihui/rmarkdown}

\hypertarget{ref-yilmaz2019whose}{}
Yilmaz, T., \& Bozo, Ö. (2019). Whose gaze is more objectifying? An
experimental study of college women's state self-objectification, body
shame, negative mood, and body dissatisfaction. \emph{Mediterranean
Journal of Clinical Psychology}, \emph{7}(2).

\hypertarget{ref-young1979}{}
Young, I. M. (1979). \emph{Proceedings of the second sex---Thirty years
later: A commemorative conference on feminist theory,}

\hypertarget{ref-zurbriggen2011self}{}
Zurbriggen, E. L., Ramsey, L. R., \& Jaworski, B. K. (2011). Self-and
partner-objectification in romantic relationships: Associations with
media consumption and relationship satisfaction. \emph{Sex Roles},
\emph{64}(7-8), 449--462.

\endgroup

\end{document}