### 2020-09-29 SEM analysis supplement/SEM Proposal.Rmd
rename coursework file to sem supplement

---
title: "Objectification in Same-Sex Pairs: Proposal"
author: "Kat Kyuchukova & Asha Hinson"
date: "3/3/2019"
output: html_document
---
##Purpose
Sexual objectification, the process by which a woman’s whole being is viewed as a collection of sexualized body parts valued predominantly for commodification, is a phenomenon that undoubtedly results in a variety of negative outcomes (Bartky, 1990). Fredrickson and Roberts (1997) highlight the objectification theory as a framework that works to place female bodies within a sociocultural context with the intention of illuminating the lived experiences and risks of women who encounter sexual objectification on a daily basis, such as their physical, psychological, and social well-being. A critical consequence of being viewed by others in ways that are sexually objectifying is that, over time, individuals may be influenced to internalize an observer's perspective on self, an effect termed trait self-objectification. Thus, the purpose of this statistical study is to explore whether partners’ objectification of each other affects trait self-objectification, and the resulting feelings of authenticity. Furthermore, this project is additionally interested in the links between exposure to sexual objectification experiences, authenticity, and postinteraction outcomes including feelings of agency in romantic relationships, career aspirations, and cognitive performance, within a same-sex interaction (Quinn, Kallen, Twenge, & Fredrickson, 2006). Empirical evidence reveals that objectification manifests through inauthenticity in romantic relationships (Lamb, 2010), adverse attitudes in regard to career aspirations, and a decrease in concentration and impairment in female cognitive performance (Swim, Hyers, Cohen, & Ferguson, 2001). By using dyadic analysis, we expect to find results consistent with previous literature to illustrate that self-objectification is both a self and interpersonal process, heightened by the real-time sexual objectification of a female interaction partner.

##Data/Population

The data collected in this study sought to replicate Garcia et al. (2016), which sought to measure objectification in previously unacquainted male-female pairs. This data was given to us by Dr. Garcia, who collected data from both Smith College and Clark University. The data was ultimately merged and filtered to only include only self-identified female same-sex pairs from both schools. The two samples are often referred to below as Sample 1 (data from Clark University) and Sample 2 (data from Smith College). These participants were mostly first-year college students, with an average age of 18.85 years (SD=1.04). The sample was 48% White/European American, 9% Black/African-American, 28% Asian/Pacific Islander, 9% Latinx, and 5% mixed-race. There were eight White/White pairs and four same race racial minority pairs, for a total of twelve same-race pairs. The remaining twenty were mixed race pairs, of which fifteen were White/racial minority pairings and five were cross-racial minority group pairs. 64% of the sample identified as heterosexual, and 23% identified as gay, lesbian or bisexual. The data is organized in a pairwise fashion such that each row corresponds to one participant in the study, and includes data for themselves and their partner. We would want to generalize these findings to general interactions between women.

##Model

![](/Users/katerinakyuchukova/Downloads/SEMmodel.jpg)

All variables in this model are measured. The endogenous variables are actor/partner SSO, authenticity, cognitive performance, career aspirations and relationship agency. The exogenous variables are actor/partner trait self-objectification, and actor objectification of partner and vice versa, as well as the interaction between the two.

The variables are as follows:

#####TSO. 
TSO was assessed using the Self-Objectification Questionnaire (Fredrickson et al., 1998; Noll & Fredrickson, 1998). This asks participants to rank order functional aspects of their bodies as well as physical self-concepts. The average rank of the functional aspects is then subtracted from the rank of the physical self-concepts. As was done in Garcia et al. (2016), all measures are multiplied by -1 so that positive scores correspond to more TSO.

#####Cognitive performance. 
Cognitive performance was assessed using trigrams from the Remote Associates Task (McFarlin & Blascovich, 1984). Ten items were selected and presented to participants. For example, the correct answer for the trigram “Quack: Pond: Waddle” would be “Duck”. Participants are limited to 30 seconds. For every correct answer, 1 point is given. The mean score was 5.03. Cognitive performance was measured first in order to measure potential immediate detriments to performance (Garcia et al., 2016).

#####SOO.
To measure the participant’s objectification of their partner in the interaction, participants were asked a series of questions about the frequency of thoughts of several aspects of their partner (Garcia et al., 2016). Questions included aspects of their partner’s internal traits such as personality, friends, family, and extracurricular interests, as well as external traits such as body, appearance, clothing, and body parts. All questions were to be rated on a scale from 1 (not at all) to 7 (constantly). Objectification was measured by getting the difference between the average frequency of thought about their partner’s external traits (α=.78 for Sample 1, α=.74 for Sample 2) and frequency of thought about their partner’s internal traits (α=.70 for Sample 1, α=.78 for Sample 2). A positive score in this scale would indicate that the participant thought about their partner’s external traits more than the partner’s internal traits, and a negative score would indicate the opposite.

#####Interaction Authenticity. 
Just like Garcia et al. (2016), we assessed the extent to which participants felt comfortable in the interaction and felt it was authentic. We did this by asking them to rate the extent to which they felt the interaction was comfortable, happy, friendly, warm, easygoing, sincere, and authentic on a scale from 1 (not at all) to 7 (very much). Participants were also asked whether they thought their partner was authentic in the interaction, and if they felt they were authentic in the interaction. These questions were ranked on a scale from 1 (not authentic at all). These were combined to form the authenticity scale (α=.91).

#####SSO. 
To assess state self-objectification, we used an average of two items from Saguy et al. (2010) that was also used in Garcia et al. (2016). Participants were asked to rank how much they agreed with the following statements: “During the interaction I felt more like a body than a full self” and “I felt more like a body than as a real person in the interaction”. Saguy et al. (2010) originally used 3 items, but the 3rd item had a low reliability score (r=*), so we chose to only use the first two for our measure of SSO.

#####Relationship Agency.
To assess how much relationship agency participants felt they would have in future relationships, we used the scale from Garcia et al. (2016), asking how likely it was that they would do the following: ‘‘ask someone out on a date,’’ ‘‘open the door for your date,’’ ‘‘pay for a date,’’ ‘‘ask your boyfriend/girlfriend to marry you,’’ ‘‘initiate sex with your girlfriend/boyfriend,’’ ‘‘initiate condom use during sex,’’ ‘‘surprise your boyfriend/ girlfriend with a gift,’’ and ‘‘ask your girlfriend/boyfriend to move with you to a new place.’’ Responses were measured on a scale ranging from 1 (not at all likely) to 7 (extremely likely). This item had high reliability for Sample 1 (α=0.73) but low reliability for Sample 2 (α=0.58).

#####Career Aspirations.
To measure career aspirations, we used the 10-item adaptation of Gray & O’Brien (2007)’s Career Aspiration Scale used in Garcia et al. (2016). This asked participants how true statements were about their future careers on a scale from 0 (not at all true of me) to 4 (very true of me). Items include ‘‘I hope to become a leader in my career field’’ and ‘‘I hope to move up through any organization or business I work in.’’ Items were fairly reliable (α=0.72 for Sample 1, α=0.81 for Sample 2).