### 2020-11-30 Final_Manuscript.tex
final edits

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Interpersonal Objectification among Women: Evidence for the Relationship between Being Objectified and Self-Objectification in Same-Gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Laneé Jung},
            pdfkeywords={Interpersonal sexual objectification, self-objectifiation, authenticity,
actual interactions, psychology of women, dyadic data analysis},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Interpersonal Objectification among Women: Evidence for the Relationship
between Being Objectified and Self-Objectification in Same-Gender
Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1, 2}, Asha Hinson\textsuperscript{1}, \&
Laneé Jung\textsuperscript{2}}
    \date{}

\shorttitle{Objectification among Women}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Smith College\\\textsuperscript{2} Program in Statistical and Data Sciences, Smith College}
\keywords{Interpersonal sexual objectification, self-objectifiation, authenticity, actual interactions, psychology of women, dyadic data analysis}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\authornote{Randi L. Garcia is in the Psychology Department
and the Program in Statistical and Data Sciences at Smith College.

The data that support the findings of this study are available on
request from the corresponding author. The data are not publicly
available due to privacy or ethical restrictions.

Correspondence concerning this article should be addressed to Randi L.
Garcia, 415 Bass Hall, Smith College, Northampton, MA 01060. E-mail:
\href{mailto:rgarcia@smith.edu}{\nolinkurl{rgarcia@smith.edu}}}

\abstract{
Existing research on interpersonal objectification has focused mostly on
links between sexual objectification, self-objectification, and negative
outcomes for women within mixed-gender interactions (Garcia, Earnshaw,
\& Quinn, 2016; Gervais, Sáez, Riemer, \& Klein, 2020). The purpose of
the present study was to extend Garcia et al. (2016)`s past research on
mixed-gender interpersonal objectification to interactions between
women. Women were brought into the laboratory and interacted
face-to-face in same-sex dyads. Dyadic analysis (Kenny, Kashy, \& Cook,
2006) was utilized to detect whether partners' objectification of each
other affected state self-objectification, as well as the resulting
feelings of comfort and authenticity during the interaction. Results
revealed a significant positive relationship between being objectified
by a woman interaction partner and a woman's own self-objectification (a
partner effect). There was no significant relationship between
self-objectification and interaction inauthenticity, but there were
significant negative effects of inauthenticity on career aspirations and
relationship agency (actor effects). The significant partner effect of
objectification on self-objectification suggests that women being
objectified by other women may result in feelings of
self-objectification, and this finding opens up questions for future
research about how interpersonal objectification might function in
situations without an immediate male gaze.

}

\begin{document}
\maketitle

Fredrickson and Roberts (1997)'s Objectification Theory suggests that,
in addition to being steeped in a culture that objectifies women, women
are objectified in actual interpersonal encounters. The negative effects
of this interpersonal objectification for women have been theorized to
be the strongest under the \enquote{male gaze,} that is, when it is a
perceived or actual man doing the objectifying (Calogero, 2004; Gay \&
Castano, 2010; Gervais, Vescio, \& Allen, 2011). What are the
consequences when a woman objectifies another woman in an interaction?
Psychological researchers studying the sexual objectification of women
have in recent years explored the interpersonal process of the
objectification, finding evidence for \emph{self}-objectification in the
target as a proximal consequence of being objectified (Garcia et al.,
2016; Riemer, Sáez, Brock, \& Gervais, 2020; Strelan \& Pagoudis, 2018).
However, this research has largely focused on mixed-gender interactions
only. In a recent review, Gervais et al. (2020) organized the extant
literature on interpersonal self-objectification and proposed a
theoretical model called the Social Interaction Model of Objectification
(SIMO). The SIMO focuses on understanding mixed-gender interactions,
acknowledging the patriarchal power structure embedded in these mixed
interactions, but can this model be extended to interactions among
women? Although there is evidence that women can objectify other women
(Gervais, Holland, \& Dodd, 2013; Loughnan et al., 2015; Puvia \& Vaes,
2013), studies investigating the process of interpersonal
objectification in interactions between women is scarce.

The current study uses the face-to-face interaction methodology deployed
in Garcia et al. (2016) to begin to answer a series of research
questions about interpersonal objectification among women: Is
objectification by a female interaction partner related to an increase
in self-objectification for the woman being objectified? Does
objectification by a woman have the same downstream negative
consequences for women as being objectified by a male interaction
partner? The current study addresses these questions by investigating
interpersonal objectification among interacting pairs where both
partners identify as women.

\subsection{Self-Objectification}\label{self-objectification}

Self-objectification is a psychological process that translates the
experiences of interpersonal objectification (Gervais et al., 2020;
Loughnan, Baldissarri, Spaccatini, \& Elder, 2017) into negative mental
health outcomes (e.g., interaction inauthenticity, anxiety, lower
self-esteem, and poor cognitive performance; Calogero, Tantleff-Dunn, \&
Thompson, 2011; Moradi \& Huang, 2008). Calogero et al. (2011) proposes
that self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one adopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (B. L. Fredrickson, Roberts,
Noll, Quinn, \& Twenge, 1998), or during video chat when you believe
your interaction partner can only see your body and not your face
(Saguy, Quinn, Dovidio, \& Pratto, 2010). This momentary
self-objectification is referred to as \emph{state} self-objectification
(SSO; Calogero et al., 2011; Moradi \& Huang, 2008) and is characterized
by feeling like a body rather than a full self within a particular
moment, instance, or context.

Perhaps the most adverse negative consequence of interpersonal
objectification is that being objectified socializes girls and women to
routinely treat themselves as objects to be looked at and evaluated
(Bartky, 1990). There is experimental and observational evidence that
being objectified by another person during an interaction elicits SSO.
For example, Loughnan et al. (2017) found that for women, imagining a
time when they were objectified by another person caused reductions in
human traits attributed to the self, and notably, the gender of the
person doing the objectification was not a moderating factor. In
addition, in a mixed-gender dyadic study of actual interactions, Garcia
et al. (2016) found that men's reported objectification of their female
interaction partner was associated with increased self-objectification
in the female partner. Further, this negative effect of the men's
objectification on women's SSO was strongest for women higher in TSO.
The current study uses this same dyadic paradigm to study same-gender,
woman-woman interacting pairs.

\subsection{Women Objectifying Women}\label{women-objectifying-women}

Past research investigating women objectifying other women has focused
on the psychological conditions that lead women to objectify other
women. Harsey and Zurbriggen (2020) found that (trait)
self-objectification was related to the objectification of women to a
similar degree for male and female participants. Along the same lines,
Strelan and Hargreaves (2005) found that the more women self-objectify,
the more they objectify other women. Could the process work in the other
direction as well? That is, could objectification by a woman produce
(state) self-objectification, in a similar manner to objectification by
a man? Hill and Fischer (2008) assessed women's experiences of
objectification from men independently from women's experiences of
objectification from other women. They found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well. This process, however, was not confirmed
to occur during an actual interpersonal interaction. Thus, there is
evidence that women do objectify other women, indeed, Loughnan et al.
(2017) found that women objectify other women to a greater extent than
they objectify men. Women have also been found to objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but very little is known about interpersonal sexual
objectification among women and what the effects of this objectification
might be. Past studies have found that objectification has more adverse
consequences for women than men (Gervais et al., 2011; Moradi \& Huang,
2008; Saguy et al., 2010), however, we do not know much about the effect
of the gender of the \emph{objectifier} on these potential detrimental
outcomes.

\subsection{Interpersonal
Objectification}\label{interpersonal-objectification}

Studies have shown that within social encounters women are gazed at more
than men (Henley, 1977), often times feel \enquote{looked at} within
interpersonal interactions (Argyle \& Williams, 1969), and are likely to
internalize the objectifying gaze on their physical self (Puvia \& Vaes,
2013). As mentioned above, interpersonal objectification has been found
to have negative consequences for women. Gervais et al. (2011) found
that an objectifying gaze by a male interaction partner (confederate)
was associated with lower math performance for women than being
objectified by a female interaction partner. Gervais et al. (2020)
recently reviewed the research on interpersonal objectification and
organized our current theoretical understanding of this process in the
SIMO. What is clear from Gervais et al. (2020)'s review is that we know
little about women-on-women interpersonal objectification, in
unstructured contexts.

Past research has found that the experience of state
self-objectification in mixed-gender contexts (Meltzer, 2020; stranger
and romantic; Strelan \& Pagoudis, 2018; Zurbriggen, Ramsey, \&
Jaworski, 2011) has negative consequences for women, but what about in
the context of interactions with other women? Is the
self-objectification experienced in mixed-gender interactions associated
with the same negative process as that experienced in same-gender
interactions? The ample research demonstrating that the male gaze has a
particularly detrimental effect for women would suggest no (Calogero,
2004; Gay \& Castano, 2010; Gervais et al., 2011; Saguy et al., 2010).
However, there is evidence that women do objectify other women (Harsey
\& Zurbriggen, 2020; Loughnan et al., 2015; Puvia \& Vaes, 2013) and
recent experimental evidence suggests that the female gaze causes no
less self-objectification than the male gaze (Yilmaz \& Bozo, 2019). It
would be helpful to understand the consequences of this
\emph{intra}group objectification for women. When women objectify other
women, does it lead to self-objectification in the same way men's
objectification of women does (Garcia et al., 2016)?

If so, does the self-objectification experienced in these same-gender
interactions have the same negative consequences for authenticity in
that interaction?

One mechanism by which interpersonal objectification translates into
negative consequences is interaction inauthenticity - when actors feel
that they are not being their true selves in an interaction. There is
some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán, Jiao, \& Aubrey,
2020). A recent study has investigated the effect of
self-objectification on the reduction of relationship building skills in
general, including same-sex friendships (Terán et al., 2020).
Authenticity reduction and interaction quality disruptions have also
been found in research on stigmatized-stigmatizer interactions (M. R.
Hebl \& Dovidio, 2005; Pearson et al., 2008; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), and we might view the experience
of being objectified in an interaction as a potential identity threat
situation (Gervais et al., 2020; Nadal \& Haynes, 2012). Further,
empirical evidence reveals that healthy relationship functioning
manifests through authenticity in romantic relationships (Brunell et
al., 2010), at least in an intergroup context such as heterosexual
romantic relationships. Evidence for other adverse consequences of
interpersonal objectification include reductions in career aspirations
(Garcia et al., 2016) and a decrease in concentration and impairment in
female cognitive performance (Kahalon, Shnabel, \& Becker, 2018; D. M.
Quinn, Chaudoir, \& Kallen, 2011). These are also negative consequences
found for women under stereotype threat (Davies, Spencer, \& Steele,
2005; Spencer, Steele, \& Quinn, 1999). When a woman is objectified by a
man, and subsequently experiences self-objectification, the intergroup
nature and power differential of this encounter might trigger threat.
Thus, me might expect there to fewer negative consequences when a woman
is objectified by another woman. That is, the intergroup threat
literature might predict that woman-woman interpersonal objectification
processes diverge from mixed-gender interpersonal objectification
processes precisely because they are not intergroup interactions, at
least with respect to gender identity.

In summary, a woman objectified by another woman may not be having the
same negative consequences that cascade from situations that trigger
group-based identity threat (Deaux \& Major, 1987; Dovidio, Hebl,
Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio, 2005); however,
feeling like a body rather than a full human (higher SSO) in any
interaction, whether intergroup or intragroup, may be enough to reduce
women's feelings of authenticity and social competence, regardless of
the gender of the objectifier (Terán et al., 2020; Tolman, Impett,
Tracy, \& Michael, 2006).

\subsection{The Current Study}\label{the-current-study}

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are potentially objectifying
each other and where both interaction partners identify as women.
Moreover, the current study used the same face-to-face interaction
paradigm as Garcia et al. (2016) and dyadic data analysis techniques to
examine the relationships for both women simultaneously. Although the
literature on intragroup objectification among women is small, the
results are mixed, and they do not cover actual interpersonal
encounters, we expected to replicate some of the results found in Garcia
et al. (2016). Most importantly, we predicted that being objectified by
one's interaction partner would be related to state self-objectification
(SSO). We also expected that TSO would moderate this relationship,
amplifying the positive association between being objectified and SSO
for women higher in TSO. Here we would like to note that Puvia and Vaes
(2013) would alternatively predict that women's tendency to
self-objectify (TSO) leads them to objectify other women (more
precisely, to dehumanize a sexualized woman). In addition, Strelan and
Hargreaves (2005) and Harsey and Zurbriggen (2020) would also predict
the TSO to other-objectification link, further theorizing that this
relationship is mediated by the woman's own state self-objectification
(SSO). Thus, the limited but extant literature on woman-woman
objectification points to a possible alternative model from the model
tested in the current study. Where appropriate, we report results
considering the causal directions implied by this alternative mediation
model.

We also tested whether SSO would, in turn, lead to feelings of
inauthenticity. The investigation of the SSO to inauthenticity
connection is considered exploratory given the lack of support for this
connection from the literature on identity threat in \emph{intra}group
interactions but support for this connection in the objectification
literature (Garcia et al., 2016; Rollero, 2016; Terán et al., 2020).
Regardless of whether there was an association between SSO and
inauthenticity, we hypothesized that feelings of inauthencity would be
associated with reduced feelings of agency in romantic relationships,
reduced career aspirations, and reduced cognitive performance. In
summary, we expected to find a positive relationship between
other-objectification by one's partner and state self-objectification.
We also expected to find a negative relationship between state-self
objectification and interaction authenticity, and that interaction
authenticity will be positively related to cognitive performance,
relationship agency, and career aspirations.

\section{Methods}\label{methods}

\subsection{Procedure}\label{procedure}

Except for the instructions given to participants, the procedure used
was identical to that in Garcia et al. (2016). In brief, that
methodology is that each participant arrived at the laboratory and were
then led into separate cubicles to prevent any communication between the
participants before the interaction. In addition, each participant was
screened for prior acquaintance to confirm that they had not met prior
to the study. They were asked to sign the consent form to participate,
and the study was described as follows: \enquote{This is a study looking
at how students form different types of relationships at college.} A
prompt on the computer screen told the participants that they were
assigned to the \enquote{College Relationships} condition and gave the
following instructions:

\begin{quote}
There are many types of relationships people form in college. During the
interaction, please think about your partner's potential as a romantic
partner. \textbf{Even if they are not the gender you are attracted to},
you can still judge their potential as a romantic partner. After the
interaction you will be asked to evaluate how dateable your partner is.
In other words, we would like to know if you think someone would date
your interaction partner. Also, your interaction partner will be
evaluating you in the same manner.
\end{quote}

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Western culture and beyond, women need to
look attractive to obtain and maintain successful relationships, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask all women participating in the study,
even heterosexual women, to judge their fellow-woman partners for their
potential as romantic partners. We felt that this prompt would keep the
study closest to a replication (and extension) of the previous Garcia et
al. (2016) version of the study, and strengthen the potential for
objectification that would normally be low in the context of a
psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners, specifically, their sexual attractiveness (Puvia
\& Vaes, 2013). Women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspected that this habitual
thought pattern would translate to their thoughts about other women.
Analyses comparing levels of other-objectification between heterosexual
and non-heterosexual women are provided in the measures section below.

After receiving the \enquote{College Relationships} prompt, the two
participants were then brought into a larger interaction room where they
sat on stools prearranged to be approximately 1 meter apart. The
experimenter instructed the participants to \enquote{get to know each
other} for 10 minutes and then left the room. After 10 minutes, the
experimenter came back into the room and stopped the interaction. The
participants then went back to their individual cubicles and completed a
set of post-interaction measures. Participants were then thanked for
their participation and debriefed. More detail on this methodology can
be found in Garcia et al. (2016).

\subsection{Participants}\label{participants}

Thirty-two previously unacquainted dyads of self-identifying women
participated in this study. Data from two demographically similar higher
education institutions in the Northeast United States were combined to
create the final analysis sample (\emph{N =} 64) used in this study. In
the measures section that follows we refer to them as Sample 1 and
Sample 2. Sample 1 (\emph{N =} 24, or 12 pairs) is from a co-ed liberal
arts college and Sample 2 (\emph{N =} 40, or 20 pairs) is from a women's
liberal arts college. Initially, data was collected from both
same-gender and mixed-gender dyads at both institutions. Sample 1
originally consisted of 22 pairs, 12 men and 32 women. In Sample 2 there
were 23 pairs made up of 43 women and one man, as well as two
participants who did not identify as either a woman or man. To
investigate only same-gender pairs of women, we limited participant data
to women in both samples. Due to difficulties in the logistics of dyadic
interaction studies, data collection was discontinued after one year at
the second institution and the sample size was taken as the number of
women-women dyads at this point in time. All results presented below are
from models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were different racial minority group pairs.
64.06\% of the sample identified as heterosexual, and 25\% identified as
gay, lesbian or bisexual. See the analysis of differences between
heterosexual and non-heterosexual women in levels of
other-objectification of their female partners in the measures section
below.

\subsection{Measures}\label{measures}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:corrtable}Correlations and Descriptive Statistics among Study Variables}
\begin{tabular}{llllllllll}
\toprule
 & \multicolumn{1}{c}{$M$} & \multicolumn{1}{c}{$SD$} & \multicolumn{1}{c}{1.} & \multicolumn{1}{c}{2.} & \multicolumn{1}{c}{3.} & \multicolumn{1}{c}{4.} & \multicolumn{1}{c}{5.} & \multicolumn{1}{c}{6.} & \multicolumn{1}{c}{7.}\\
\midrule
1. Actor's trait self objectification (TSO) & -0.35 & 2.64 &  &  &  &  &  &  & \\
2. Actor's objectification of partner & -1.58 & 1.21 & .20 &  &  &  &  &  & \\
3. Partner's objectification of the actor & -1.58 & 1.21 & .09 & .14 &  &  &  &  & \\
4. Actor's state self-objectification & 1.92 & 1.13 & .13 & -.09 & .30* &  &  &  & \\
5. Actor's authenticity of interaction & 5.23 & 1.02 & -.02 & -.07 & .03 & -.10 &  &  & \\
6. Actor's future relationship agency & 4.69 & 0.96 & .04 & .09 & -.10 & -.09 & .23+ &  & \\
7. Actor's cognitive performance & 5.03 & 2.29 & .08 & .11 & .06 & .02 & .11 & .07 & \\
8. Actor's career aspirations & 3.88 & 0.61 & .05 & -.14 & -.08 & .01 & .33** & .26* & -.07\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} * p < .05. ** p < .01.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{lltable}

\small{
\begin{longtable}{lllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:measures}Measures}\\
\toprule
 & \multicolumn{1}{c}{Original Source} & \multicolumn{1}{c}{Example Item} & \multicolumn{1}{c}{Reliability Sample 1} & \multicolumn{1}{c}{Reliability Sample 2}\\
\midrule
Cognitive performance & @mcfarlin1984remote & The correct answer for the trigram "Quack: Pond: Waddle" would be "Duck" & Not avaliable & Not avaliable\\
Objectification of partner & @garcia2016objectification & "During the interaction, how often did you think about your partner’s personality?" (...friends, family, and extracurricular interests, body, appearance, clothing, and body parts) on a scale from 1 (not at all) to 7 (constantly). A difference between the average internal and exxternal traits is taken & external:  0.79 ; internal:  0.71 & external:  0.76 ; internal:  0.79\\
Authenticity of interaction & @garcia2016objectification & "During the interaction, to what extent did you feel comfortable?" (...happy, friendly, warm, easygoing, sincere, and authentic) on a scale ranging from 1 (not at all) to 7 (very much) & 0.91 & 0.91\\
State self-objectification & @saguyetal2010 & "During the interaction I felt more like a body than a full self" & 0.84 & 0.85\\
Relationship agency & @garcia2016objectification & "ask someone out on a date" Responses were measured on a scale ranging from 1 (not at all likely) to 7 (extremely likely) & 0.72 & 0.74\\
Career aspirations & @grayobrien2007 & "I hope to become a leader in my career field" & 0.73 & 0.80\\
Trait self objectification (TSO) & @nollfredrickson1998 & Participants rank order aspects of their bodies from 1 (least important) to 10 (most important): weight, sex appeal, physical attractiveness, firm/sculpted muscles and body measurements, strength, physical coordination, energy level, health and physical fitness. Difference scores were computed by subtracting the sum of the 5 functional aspects/competence attributes (e.g., health, strength) from the sum of the 5 physical self-concepts/appearance attributes (e.g., physical attractiveness, weight), and all measures were multiplied by -1 so that positive scores indicated greater TSO. & appearance:  0.44 ; competence:  0.43 & appearance:  0.51 ; competence:  0.30\\
\bottomrule
\end{longtable}
}
\end{lltable}

The same set of measures given in Garcia et al. (2016) were given to
participants in the current study. For more details about each of these
measures please refer to that previous paper. To summarize here, the
list of measures in the order given to participants appears in
Table~\ref{tab:measures} along side the original source for each scale,
an example item, and the reliability in the current sample. Correlations
and descriptive statistics of all study variables appear in
Table~\ref{tab:corrtable}.

Cognitive performance as measured with the Remote Associates Task's
trigrams (McFarlin \& Blascovich, 1984) was given first after the
interaction to capture any immediate cognitive effects of the
interaction. Next participants were given a set of self-report measures
about their thoughts and feelings during the interaction that included
State-Other Objectification, Interaction Authenticity, and State
Self-Objectification. Breaking from the Garcia et al. (2016)
methodology, we used a three item state self-objectification scale as
originally given in Saguy et al. (2010), but in both samples the
reliability of the scale was higher once the third item was removed, so
we chose to only use the first two items for our measure of SSO.
Following these interaction measures, participants completed the
Relationship Agency Scale and the Career Aspirations Scale. The
Relationship agency scale originally had 9 items, but the 9th item had
low correlations with the remaining items so it was trimmed. As a
result, the 8-item scale had moderately high reliability for both
samples (see Table~\ref{tab:measures}). Lastly, as was the case in
Garcia et al. (2016), Trait Self-Objectification was given last, as not
to arouse suspicious at the start of the study that we were actually
focusing on objectification and to give enough time that immediate
effects of the interaction would have worn off.

As can be seen in Table~\ref{tab:corrtable}, the mean
other-objectification of women by women was \emph{M =} -1.58 (\emph{SD
=} 1.21). This corresponds to women objectifying other women to a
\emph{greater} extent than women's objectification of men reported in
Garcia et al. (2016) (\emph{M =} -1.68, \emph{SD =} 1.52). Further, in
the current sample the difference in other-objectification between
heterosexual (\emph{M =} -1.77, \emph{SD =} 1.14) and non-heterosexual
women (\emph{M =} -1.39, \emph{SD =} 1.35) was not statistically
significant, \emph{t}(25.89) = 1.02, \emph{p =} 0.32.

\section{Results}\label{results}

We used R (Version 3.5.2; R Core Team, 2017) and the R-packages
\emph{tidyverse} (Version 1.2.1; Wickham, 2017), \emph{psych} (Version
1.8.4; Revelle, 2017), and \emph{nlme} (Version 3.1.137; Pinheiro,
Bates, DebRoy, Sarkar, \& R Core Team, 2017) for our analyses. Further,
we used the R-packages \emph{papaja} (Version 0.1.0.9842; Aust \& Barth,
2018), \emph{apaTables} (Version 2.0.5; Stanley, 2018), with
\emph{knitr} (Version 1.25; Xie, 2015) to create a fully reproducible
manuscript in \emph{rmarkdown} (Version 1.15; Xie, Allaire, \&
Grolemund, 2018).

The R-package \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018)
was used to create a fully reproducible APA style manuscript containing
the analysis code and manuscript text integrated into one source file.
The benefit of this integration is that the numbers reported throughout
this document are coded into the text, ensuring that no errors were made
during a workflow marked by copying and pasting from statistical
software to word processing software. Note that here we are using the
term reproducibility to mean getting the same results when running
analyses again using the \emph{same} data, not to be confused with
replicability meaning collecting a new dataset using the same methods
and obtaining the same results (Patil, Peng, \& Leek, 2016). The source
code for this manuscript along with survey materials and experimenter
scripts for the study protocol are all available at
\url{https://github.com/kkyuchukova/object-in-action}. The data
unfortunately cannot be made publicly available given the dyadic nature
of the observations. That is, with dyadic data, if a person who
participated in the study found their own scores, using the dyad
identification number, they could then see their partner's scores and
thus confidentiality would be breached.

\subsection{Analysis Strategy}\label{analysis-strategy}

\begin{figure}
\includegraphics[width=400px]{figures/APIM_figure} \caption{Basic actor-partner interdependence model (APIM) depiction.}\label{fig:apim}
\end{figure}

This study sought to detect whether partners' objectification of one
another affected state self-objectification (SSO). Specifically, we were
interested in testing the relationship between state-other
objectification and SSO, and how SSO in turn, may or may not affect
feelings of inauthenticity during the interaction. In addition, we also
tested if the effect of other-objectification in an interaction on SSO
is only present for those women who are high in trait
self-objectification (moderation effect). Further, we investigated the
relationships between experiencing interaction inauthenticity and
relationship agency, career aspirations, and cognitive performance.

While past studies investigating objectification in interactions used
dyadic path analysis (Garcia et al., 2016), the current study used
multilevel modeling procedures. Dyadic analyses for distinguishable
dyads (e.g., mixed-gender interacting pairs) is more natural using
Structural Equation Modeling (SEM) than it is for indistinguishable
dyads (e.g., same-gender interacting pairs), as we had in the current
study (Garcia, Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One
reason for this asymmetry is that, due to the arbitrary distinctions
made between \enquote{partner 1} and \enquote{partner 2} in
indistinguishable dyads, many estimates need to be fixed to be equal
(i.e., paths, variances, covariances, endogenous intercepts, and
exogenous means), but these equality constraints should not then be
considered in the degrees of freedom calculations for fit estimations
(Olsen \& Kenny, 2006). Further, Olsen and Kenny (2006) details how a
new independence model and the corresponding fit measure should be
re-calculated for indistinguishable dyads models. The current study used
dyadic multilevel modeling (MLM) to test all relationships, moderation,
and mediation patterns. The online supplementary materials found at
\url{https://github.com/kkyuchukova/object-in-action} contain model
estimates obtained using SEM. See Ledermann and Kenny (2017) for a more
complete discussion of the considerations for using SEM versus MLM for
dyadic analysis.

Testing hypotheses and exploring relationships in the current sample of
indistinguishable dyads involved using the Actor-Partner Independence
Model (APIM) approach for each outcome variable. See
Figure~\ref{fig:apim} for a basic APIM model. The APIM estimates effects
due to one's own, as well as one's partner's, predictor variables on the
one's own outcome variable. The current study deals with
indistinguishable dyads, meaning the designation of who is the
\enquote{actor} and who is the \enquote{partner} is arbitrary. In total
we ran five APIM's---one for each outcome variable---to test the series
of the relationships proposed (i.e., for SSO, inauthenticity, career
aspirations, relationship agency, and cognitive performance).

\subsection{Main Results}\label{main-results}

\begin{figure}
\includegraphics[width=400px]{figures/SEMfigure} \caption{This figure depicts the relationships tested bewteen study variables. The light gray variables represent redundant variables and are shown to emphasize the dyadic nature of the data. The estimates were from obtained from separate MLMs. Sample was controlled for in all models. The effects of TSO, actor other-objectification, and partner other-objectification are from a model where the interaction of TSO and partner other-objectification had been trimmed. *p < .05, dashed lines represent non-significant relationships.}\label{fig:semfigure}
\end{figure}

All relationships between study variables and the MLM estimates are
depicted in Figure~\ref{fig:semfigure}.

First, we used the APIM to test for evidence of a partner effect of
other objectification and SSO in this sample of same-gender woman-woman
interacting dyads. In Garcia et al. (2016), men's objectification of
women was significantly related to women's SSO, the current study sought
to replicate this partner effect. As expected, the partner effect of
other-objectification on SSO in the current all-women sample was
statistically significant, \emph{b} = 0.29, \emph{SE} = 0.12, \emph{p} =
.019, \emph{95\%CI} = {[}0.06, 0.53{]}. That is, the extent to which
one's partner reported thinking about a woman's external characteristics
more than her internal characteristics was significantly related to the
woman's own reported feeling more like a body than a full self. Along
side the Garcia et al. (2016) study, there is now evidence that
interpersonal objectification in an actual interpersonal encounter is
related to state self-objectification for women both when the
objectifier is a man and when the objectifier is a woman.

In addition to partner other-objectification, this model also included
one's own objectification of their partner (actor
other-objectification), actor trait self-objectification (TSO), and the
interaction of partner other-objectification and TSO. One's own other
objectification had no significant effect on SSO (actor effect of
other-objectification), \emph{b} = -0.16, \emph{SE} = 0.12, \emph{p} =
.210, \emph{95\%CI} = {[}-0.4, 0.09{]}. Further, inconsistent with past
findings, there was no statistically significant interaction of
partner's other objectification and the person's trait
self-objectification on SSO, \emph{b} = 0.03, \emph{SE} = 0.27, \emph{p}
= .910, \emph{95\%CI} = {[}-0.5, 0.56{]}. There was also no significant
main effect of trait self-objectification on SSO \emph{b} = 0.07,
\emph{SE} = 0.05, \emph{p} = .183, \emph{95\%CI} = {[}-0.03, 0.17{]}.
The \(R^2\) of this model (which includes the partner effect of
other-objectification discussed in the previous paragraph) was only
0.03.

As a test of the potential alternative models discussed in the
introduction---if trait-self-objectification is positively related to
objectifying one's partner---we ran a MLM with other-objectification
predicted by actor TSO. There was no statistically significant
relationship between these two variables, \emph{b} = 0.09, \emph{SE} =
0.06, \emph{p} = .133, \emph{95\%CI} = {[}-0.03, 0.2{]}. We also ran a
model with both TSO and SSO predicting other-objectification, given past
theorizing that state self-objectification might mediate the
relationship between TSO and objectification of fellow women, but there
was also no effect of SSO on other-objectification in this model,
\emph{b} = -0.17, \emph{SE} = 0.13, \emph{p} = .199, \emph{95\%CI} =
{[}-0.42, 0.08{]}.

We next moved on to test the link between SSO and interaction
authenticity. There was no significant effect of SSO on interaction
authenticity, although the estimate of this effect was in the
hypothesized negative direction, \emph{b} = -0.09, \emph{SE} = 0.12,
\emph{p} = .431, \emph{95\%CI} = {[}-0.32, 0.14{]}. Because authenticity
was a composite score of 9 items, two of which were interaction specific
authenticity items, we also estimated the pairwise correlations between
SSO and all these items individually. They were all small, ranging from
only -0.01 to -0.14. Although we hypothesized that SSO would mediate the
relationship between partner's other objectification and interaction
authenticity, after finding no relationship between SSO and
authenticity, we also tested if the partner's other objectification had
a direct effect on authenticity, but this effect was not significant,
\emph{b} = 0.06, \emph{SE} = 0.12, \emph{p} = .608, \emph{95\%CI} =
{[}-0.17, 0.29{]} (nor was the total effect of partner's other
objectification on authenticity, \emph{b} = 0.03, \emph{SE} = 0.11,
\emph{p} = .787, \emph{95\%CI} = {[}-0.19, 0.25{]}). Note that these
analyses of the relationship between SSO and inauthenticity in this
same-gender sample were considered exploratory, given that prior
research on intragroup interactions points to mixed possibilities.

Lastly, although there was no evidence that SSO was related to
interaction authenticity in the current sample, we tested if interaction
authenticity (composite of nine items) had effects on cognitive
performance, career aspirations, and relationship agency, as it did in
Garcia et al. (2016). We again used MLM and thus, these effects were
tested in three separate multilevel linear models. There was no
significant effect of interaction authenticity on cognitive performance,
\emph{b} = 0.32, \emph{SE} = 0.28, \emph{p} = .258, \emph{95\%CI} =
{[}-0.23, 0.86{]}, but authenticity was significantly positively related
to both career aspirations, \emph{b} = 0.18, \emph{SE} = 0.07, \emph{p}
= .010, \emph{95\%CI} = {[}0.05, 0.32{]} (\(R^2 =\) 0.04), and
relationship agency, \emph{b} = 0.23, \emph{SE} = 0.12, \emph{p} = .049,
\emph{95\%CI} = {[}0.01, 0.46{]} (\(R^2 =\) 0.01).

\section{Discussion}\label{discussion}

The current study tested whether the model of interpersonal
objectification and state self-objectification (SSO) used in Garcia et
al. (2016) replicates in a sample of women engaging in actual dyadic
interactions with each other. Although past research has found that
women do objectify other women (Harsey \& Zurbriggen, 2020; Loughnan et
al., 2015; Puvia \& Vaes, 2013), this is the first study to test if
interpersonal other-objectification by women during \emph{actual}
interactions is related to state self-objectification in their
woman-identified interaction partners. As hypothesized, the current
study did find a significant relationship between a woman's interaction
partner's report of having objectified her and her own post-interaction
feelings of self-objectification. That is, there was a significant
partner effect of other-objectification on SSO. This effect extends the
equivalent relationship found in mixed-gender interactions to the
context of same-gender interactions between women. Thus, evidence
suggests that it is not only the real or imagined male gaze that is
related to women's state self-objectification, but there is now also
evidence that being objectified by another woman could be related to
women's SSO, at least in the context of a scenario where they know they
are being evaluated as a potential dating partner.

As is the case in all correlational studies, we cannot be sure about the
causal direction between other-objectification and SSO. The linear model
used by the current study implies that being objectified by a woman
leads women to self-objectify, but it could be that women's SSO causes
them to be objectified by their interaction partner. This latter
interpretation is possible given the empirical evidence that it is
\emph{women's} state self-objectification that relates to being
objectified by one's partner---men-identifying partners (Garcia et al.,
2016) and now women-identifying partners. Objectification Theory
(Fredrickson \& Roberts, 1997), as well as some past experimental
studies (for example, Saguy et al., 2010), suggest that the causal flow
is from other-objectification to SSO. However, some studies
investigating women objectifying other women (Puvia \& Vaes, 2013), have
found evidence that women's \emph{trait} self-objectification is related
to women objectifying (dehumanizing) other women and this link is
mediated by \emph{state} self-objectification. The process of
interpersonal objectification among women could also very well contain a
feedback loop. That is, perhaps we tend to objectify other women who are
objectifying us. Then, if there indeed tends to be reciprocity in sexual
objectification among women, a vicious cycle could ensue through a
process of TSO causing SSO which in turn causes other-objectification of
one's partner, which in turn causes one's partner to objectify us, and
finally, being objectified further increases our own SSO. However, the
current study does not provide any evidence of the TSO to
other-objectification link (through SSO or directly)---TSO was not
significantly correlated with SSO. Further, actor's
other-objectification was not correlated with partner's
other-objectification in the current sample. Future experimental work
testing if sexual objectification by a woman can cause an increase in a
woman's state self-objectification would be helpful.

Where the results of the current study diverge most notably from the
results of studies testing interpersonal objectification among
mixed-gender dyads is the lack of evidence for relationships between SSO
and interaction inauthenticity (and, in turn, cognitive functioning).
This is somewhat surprising given the extant evidence linking SSO and
cognitive functioning (B. L. Fredrickson et al., 1998; and see Moradi \&
Huang, 2008 for a review; Quinn, Chaudoir, \& Kallen, 2011) and the
research on interpersonal other-objectification and cognitive
functioning (Garcia et al., 2016; Gervais et al., 2011; Logel et al.,
2009). This lack of evidence could potentially signal diverging
processes between women's experiences with interpersonal objectification
from men and interpersonal objectification from women. There is quite a
bit of evidence suggesting that the male gaze is particularly
detrimental (Calogero, 2004; Fredrickson \& Roberts, 1997; Gervais et
al., 2011) for women, and perhaps the self-objectification experienced
within an interaction with a women is qualitatively different from the
state self-objectification experienced within an interaction with a man.
However, as a strong note of caution, we cannot interpret a null result
as evidence of no relationship and more research on interpersonal
objectification among women is needed.

The lack of evidence for a relationship between SSO and interaction
authenticity is surprising and again, should not be interpreted as
evidence of no relationship. The estimate of this relationship was small
(close to zero), but in the negative direction. If perhaps it is the
case that there is a smaller (i.e., weaker) or zero connection between
women's feelings of SSO and inauthenticity in interactions with other
women than in interactions with men, models of interpersonal
objectification, like the SIMO (Gervais et al., 2020), could be extended
by including gender of the objectifier/interaction partner as a
moderator. Inauthenticity could be added as a potential moderated
mediating factor to help understand the circumstances that require
other-objectification and SSO to have negative consequences for women.
Perhaps one important difference is the lack of a power differential
across gendered lines---the patriarchal culture is present, but the
interaction partners are not a stigmatizer-stigmatized pair in
interpersonal objectification encounters among women. Lastly, This
theoretical work on intragroup sexual objectification among women could
potentially benefit from making connections to the small but growing
literature investigating same-race racial minority interactions (Garcia,
Bergsieker, \& Shelton, 2017; Taylor, Garcia, Shelton, \& Yantis, 2018).

Although the current study did not find a connection between SSO and
authenticity, we did find significant positive relationships between
authenticity and relationship agency, and authenticity and career
aspirations. The relationship between authenticity and cognitive
functioning was estimated as positive (the point estimate), but was not
statistically significant. Again, due to the lack of connection between
SSO and authenticity, we found no evidence of \emph{indirect}
relationships between SSO and these outcome variables. This evidence of
a relationship between authenticity and the relational outcome variables
(i.e., relationship agency and career aspirations) provides evidence
that corroborates past findings that felt authenticity in interactions
is important for healthy relationship functioning (Garcia et al., 2016;
Terán et al., 2020) and is related to mental health correlates (Tolman
et al., 2006). Just as authenticity has been found to be important in
intergroup interactions (Brunell et al., 2010; Garcia et al., 2016), we
find more evidence here that disruptions in feelings of authenticity can
negatively impact relationships beyond the current interaction partner.
Although the current study did not find a connection between
authenticity and SSO, this seems to us to remain a theoretically viable
connection, and future work must explore when and how SSO leads to
inauthenticity in interpersonal objectification situations.

\subsection{Limitations and Future
Directions}\label{limitations-and-future-directions}

\subsubsection{Sample Characteristics}\label{sample-characteristics}

Although the sample size was relatively small, due to the difficulty in
collecting data on actual interpersonal encounters, we believe that the
results from the current analyses are important and push work in
interpersonal objectification forward in an important way. Of course,
more data is better, especially for evaluating null results. Follow-up
studies investigating interpersonal objectification among women are
necessary and should strive for larger samples. In addition to being
relatively small and combined across two institutions, another
limitation of the current study sample is that the it was comprised of
mostly women from Western cultures. Being that self-objectification has
been found to be most prevalent in Western culture (Loughnan et al.,
2015), research on objectification that includes people from non-Western
cultures has been scarce (Moradi \& Huang, 2008), although more recent
work has examined objectification from a cross-cultural framework
(Loughnan et al., 2015; Wollast et al., 2020).

Further, sexualizing experiences and self-objectification are thought to
begin a very young age, but researchers have only recently begun to
examine such experiences among children (e.g., Holland \& Haslam, 2016;
Jongenelis, Byrne, \& Pettigrew, 2014). The mean age of the current
sample was 18.85 years, a young adult sample, limiting the
generalizability of our results to this age group. More longitudinal
studies investigating the developmental experience of interpersonal
objectification and self-objectification are recommended, especially
because the processes of self-objectification have been found to unfold
between the periods of adolescence and young adulthood (Vangeel,
Vandenbosch, \& Eggermont, 2018).

The current sample contained a mixture of both heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential dating/romantic partner. We
think that the heterosexual women in our sample were able to do this
with other women as their target---indeed, there is evidence that they
might do this readily (Puvia \& Vaes, 2013; Strelan \& Hargreaves,
2005)---but, perhaps they may be even more apt to activate social
comparison processes (Festinger, 1954) than women who are sexually
attracted to other women (non-heterosexual women). Social comparison
processes have been linked to self-objectification and negative body
image among women in general (Lindner, Tantleff-Dunn, \& Jentsch, 2012;
Tylka \& Sabik, 2010) and, more recently, specifically with regard to
consumption of social media site content (Feltman \& Szymanski, 2018;
Fox \& Vendemia, 2016; Hanna et al., 2017). However, these none of these
social comparison studies compared heterosexual women to
non-heterosexual women. If it is the case that there are two
differential psychological processes between heterosexual and lesbian
women at play, this added variance might have served to dampen our
ability to detect the proposed relationships between study variables.

An additional issue with including both heterosexual and lesbian women
in the sample is that the interpersonal encounter in the current study
was experience as a potentially romantic context for some and as a
strictly platonic context for others. Interpersonal sexual
objectification has been studied in both romantic and platonic contexts.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen et
al., 2011). Evidence is the romantic context is mixed, suggesting that
women can sometimes have positive outcomes from appearance valuation
within the context of an established romantic relationship. There is
also evidence that they experience less SSO after appearance comments
within established romantic relationships (Meltzer, 2020). Less is known
about first dates and there is little research on objectification
processes in lesbian relationships specifically. Further,
objectification might also occur among platonic friends: Perhaps women
could be amplifying each others' state self-objectification
\emph{before} the theoretical encounters with men, and the processes
organized in the SIMO, occur.

\section{Conclusion}\label{conclusion}

To the best of our knowledge, this was the first study to test the
relationship between objectification and self-objectification within
actual interpersonal interactions among women. The results of this study
replicated the partner effect of other-objectification on state
self-objectification found in Garcia et al. (2016) in a sample of
same-gender pairs all identifying as women. That is, the more a woman's
female interaction partner reported having objectified her, the more she
reported a state of self-objectification during the interaction. Unlike
past research finding that trait self-objectification moderated this
relationship, we did not find any evidence of such an effect. Further,
we did not find that inauthenticity was an important mediating factor
between self-objectification and the negative psychological and social
consequences for women objectified by women, as was the case in past
studies on interpersonal objectification of women by men. Further
research should be conducted with larger samples that investigate
interpersonal objecticiation among women and compare these processes to
parallel processes among mixed-gender dyads.

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-argyle1969}{}
Argyle, M., \& Williams, M. (1969). Observer or observed? A reversible
perspective in person perception. \emph{Sociometry}, 396--412.

\hypertarget{ref-R-papaja}{}
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Bartky}{}
Bartky, S. (1990). \emph{Femininity and domination: Studies in the
phenomenology of oppression}. Routledge, New York.

\hypertarget{ref-brunelletal2010}{}
Brunell, A. B., Kernis, M. H., Goldman, B. M., Heppner, W., Davis, P.,
Cascio, E. V., \& Webster, G. D. (2010). Dispositional authenticity and
romantic relationship functioning. \emph{Personality and Individual
Differences}, \emph{48}(8), 900--905.
doi:\href{https://doi.org/10.1016/j.paid.2010.02.018}{10.1016/j.paid.2010.02.018}

\hypertarget{ref-calogero2004test}{}
Calogero, R. M. (2004). A test of objectification theory: The effect of
the male gaze on appearance concerns in college women. \emph{Psychology
of Women Quarterly}, \emph{28}(1), 16--21.

\hypertarget{ref-calogero2011}{}
Calogero, R. M., Tantleff-Dunn, S. E., \& Thompson, J. (2011).
\emph{Self-objectification in women: Causes, consequences, and
counteractions.} American Psychological Association.

\hypertarget{ref-davies2005clearing}{}
Davies, P. G., Spencer, S. J., \& Steele, C. M. (2005). Clearing the
air: Identity safety moderates the effects of stereotype threat on
women's leadership aspirations. \emph{Journal of Personality and Social
Psychology}, \emph{88}(2), 276.

\hypertarget{ref-deaux1987putting}{}
Deaux, K., \& Major, B. (1987). Putting gender into context: An
interactive model of gender-related behavior. \emph{Psychological
Review}, \emph{94}(3), 369.

\hypertarget{ref-dovidio2006nonverbal}{}
Dovidio, J. F., Hebl, M., Richeson, J. A., \& Shelton, J. N. (2006).
Nonverbal communication, race, and intergroup interaction.

\hypertarget{ref-feltman2018instagram}{}
Feltman, C. E., \& Szymanski, D. M. (2018). Instagram use and
self-objectification: The roles of internalization, comparison,
appearance commentary, and feminism. \emph{Sex Roles}, \emph{78}(5-6),
311--324.

\hypertarget{ref-festinger1954theory}{}
Festinger, L. (1954). A theory of social comparison processes.
\emph{Human Relations}, \emph{7}(2), 117--140.

\hypertarget{ref-fox2016selective}{}
Fox, J., \& Vendemia, M. A. (2016). Selective self-presentation and
social comparison through photographs on social networking sites.
\emph{Cyberpsychology, Behavior, and Social Networking}, \emph{19}(10),
593--600.

\hypertarget{ref-fredrickson1998swimsuit}{}
Fredrickson, B. L., Roberts, T.-A., Noll, S. M., Quinn, D. M., \&
Twenge, J. M. (1998). That swimsuit becomes you: Sex differences in
self-objectification, restrained eating, and math performance.
\emph{Journal of Personality and Social Psychology}, \emph{75}(1), 269.

\hypertarget{ref-robertsfredrickson}{}
Fredrickson, \& Roberts. (1997). Objectification theory: Toward
understanding women's lived experiences and mental health risks.
\emph{Psychology of Women Quarterly}, \emph{21}(2), 173--206.

\hypertarget{ref-garcia2017racial}{}
Garcia, R. L., Bergsieker, H. B., \& Shelton, J. N. (2017). Racial
attitude (dis) similarity and liking in same-race minority interactions.
\emph{Group Processes \& Intergroup Relations}, \emph{20}(4), 501--518.

\hypertarget{ref-garcia2016objectification}{}
Garcia, R. L., Earnshaw, V. A., \& Quinn, D. M. (2016). Objectification
in action: Self-and other-objectification in mixed-sex interpersonal
interactions. \emph{Psychology of Women Quarterly}, \emph{40}(2),
213--228.

\hypertarget{ref-garcia2015moderation}{}
Garcia, R. L., Kenny, D. A., \& Ledermann, T. (2015). Moderation in the
actor--partner interdependence model. \emph{Personal Relationships},
\emph{22}(1), 8--29.

\hypertarget{ref-gay2010my}{}
Gay, R. K., \& Castano, E. (2010). My body or my mind: The impact of
state and trait objectification on women's cognitive resources.
\emph{European Journal of Social Psychology}, \emph{40}(5), 695--703.

\hypertarget{ref-gervais2013my}{}
Gervais, S. J., Holland, A. M., \& Dodd, M. D. (2013). My eyes are up
here: The nature of the objectifying gaze toward women. \emph{Sex
Roles}, \emph{69}(11-12), 557--570.

\hypertarget{ref-gervais2020social}{}
Gervais, S. J., Sáez, G., Riemer, A. R., \& Klein, O. (2020). The social
interaction model of objectification: A process model of goal-based
objectifying exchanges between men and women. \emph{British Journal of
Social Psychology}, \emph{59}(1), 248--283.

\hypertarget{ref-gervais2011you}{}
Gervais, S. J., Vescio, T. K., \& Allen, J. (2011). When what you see is
what you get: The consequences of the objectifying gaze for women and
men. \emph{Psychology of Women Quarterly}, \emph{35}(1), 5--17.

\hypertarget{ref-hanna2017contributions}{}
Hanna, E., Ward, L. M., Seabrook, R. C., Jerald, M., Reed, L.,
Giaccardi, S., \& Lippman, J. R. (2017). Contributions of social
comparison and self-objectification in mediating associations between
facebook use and emergent adults' psychological well-being.
\emph{Cyberpsychology, Behavior, and Social Networking}, \emph{20}(3),
172--179.

\hypertarget{ref-harsey2020men}{}
Harsey, S. J., \& Zurbriggen, E. L. (2020). Men and women's
self-objectification, objectification of women, and sexist beliefs.
\emph{Self and Identity}, 1--8.

\hypertarget{ref-hebl2005promoting}{}
Hebl, M. R., \& Dovidio, J. F. (2005). Promoting the ``social'' in the
examination of social stigmas. \emph{Personality and Social Psychology
Review}, \emph{9}(2), 156--182.

\hypertarget{ref-henley1977body}{}
Henley, N. (1977). \emph{Body politics: Power, sex, and nonverbal
communication}. Prentice Hall.

\hypertarget{ref-hill2008examining}{}
Hill, M. S., \& Fischer, A. R. (2008). Examining objectification theory:
Lesbian and heterosexual women's experiences with sexual-and
self-objectification. \emph{The Counseling Psychologist}, \emph{36}(5),
745--776.

\hypertarget{ref-holland2016}{}
Holland, E., \& Haslam, N. (2016). Cute little things: The
objectification of prepubescent girls. \emph{Psychology of Women
Quarterly}, \emph{40}(1), 108--119.

\hypertarget{ref-jongenelis2014}{}
Jongenelis, M. I., Byrne, S. M., \& Pettigrew, S. (2014).
Self-objectification, body image disturbance, and eating disorder
symptoms in young australian children. \emph{Body Image}, \emph{11}(3),
290--302.

\hypertarget{ref-kahalon2018don}{}
Kahalon, R., Shnabel, N., \& Becker, J. C. (2018). ``Don't bother your
pretty little head'' appearance compliments lead to improved mood but
impaired cognitive performance. \emph{Psychology of Women Quarterly},
\emph{42}(2), 136--150.

\hypertarget{ref-kenny2006dyadic}{}
Kenny, D. A., Kashy, D. A., \& Cook, W. L. (2006). \emph{Dyadic data
analysis}. Guilford press.

\hypertarget{ref-ledermann2017analyzing}{}
Ledermann, T., \& Kenny, D. A. (2017). Analyzing dyadic data with
multilevel modeling versus structural equation modeling: A tale of two
methods. \emph{Journal of Family Psychology}, \emph{31}(4), 442.

\hypertarget{ref-lindner2012social}{}
Lindner, D., Tantleff-Dunn, S., \& Jentsch, F. (2012). Social comparison
and the ``circle of objectification''. \emph{Sex Roles}, \emph{67}(3-4),
222--235.

\hypertarget{ref-logel2009interacting}{}
Logel, C., Walton, G. M., Spencer, S. J., Iserman, E. C., Hippel, W.
von, \& Bell, A. E. (2009). Interacting with sexist men triggers social
identity threat among female engineers. \emph{Journal of Personality and
Social Psychology}, \emph{96}(6), 1089.

\hypertarget{ref-loughnan2017internalizing}{}
Loughnan, S., Baldissarri, C., Spaccatini, F., \& Elder, L. (2017).
Internalizing objectification: Objectified individuals see themselves as
less warm, competent, moral, and human. \emph{British Journal of Social
Psychology}, \emph{56}(2), 217--232.

\hypertarget{ref-loughnan2015exploring}{}
Loughnan, S., Fernandez-Campos, S., Vaes, J., Anjum, G., Aziz, M.,
Harada, C., \ldots{} Tsuchiya, K. (2015). Exploring the role of culture
in sexual objectification: A seven nations study. \emph{Revue
Internationale de Psychologie Sociale}, \emph{28}(1), 125--152.

\hypertarget{ref-mcfarlin1984remote}{}
McFarlin, D. B., \& Blascovich, J. (1984). On the remote associates test
(rat) as an alternative to illusory performance feedback: A
methodological note. \emph{Basic and Applied Social Psychology},
\emph{5}(3), 223--229.

\hypertarget{ref-meltzer2020women}{}
Meltzer, A. L. (2020). Women can benefit from sexual and physical
valuation in the context of a romantic relationship. \emph{Personality
and Social Psychology Bulletin}, \emph{46}(2), 243--257.

\hypertarget{ref-moradi2008}{}
Moradi, B., \& Huang, Y.-P. (2008). Objectification theory and
psychology of women: A decade of advances and future directions.
\emph{Psychology of Women Quarterly}, \emph{32}(4), 377--398.

\hypertarget{ref-morry2001magazine}{}
Morry, M. M., \& Staska, S. L. (2001). Magazine exposure:
Internalization, self-objectification, eating attitudes, and body
satisfaction in male and female university students. \emph{Canadian
Journal of Behavioural Science/Revue Canadienne Des Sciences Du
Comportement}, \emph{33}(4), 269.

\hypertarget{ref-nadal2012effects}{}
Nadal, K. L., \& Haynes, K. (2012). Women's psychology. women and mental
disorders. In N. Lundberg-Love P. K. \& M. A. Paludi (Eds.), (pp.
87--101). Praeger/ABC-CLIO.

\hypertarget{ref-olsen2006structural}{}
Olsen, J. A., \& Kenny, D. A. (2006). Structural equation modeling with
interchangeable dyads. \emph{Psychological Methods}, \emph{11}(2), 127.

\hypertarget{ref-patil2016statistical}{}
Patil, P., Peng, R. D., \& Leek, J. T. (2016). A statistical definition
for reproducibility and replicability. \emph{BioRxiv}, 066803.

\hypertarget{ref-pearson2008fragility}{}
Pearson, A. R., West, T. V., Dovidio, J. F., Powers, S. R., Buck, R., \&
Henning, R. (2008). The fragility of intergroup relations: Divergent
effects of delayed audiovisual feedback in intergroup and intragroup
interaction. \emph{Psychological Science}, \emph{19}(12), 1272--1279.

\hypertarget{ref-R-nlme}{}
Pinheiro, J., Bates, D., DebRoy, S., Sarkar, D., \& R Core Team. (2017).
\emph{nlme: Linear and nonlinear mixed effects models}. Retrieved from
\url{https://CRAN.R-project.org/package=nlme}

\hypertarget{ref-puvia2013being}{}
Puvia, E., \& Vaes, J. (2013). Being a body: Women's appearance related
self-views and their dehumanization of sexually objectified female
targets. \emph{Sex Roles}, \emph{68}(7-8), 484--495.

\hypertarget{ref-quinnetal}{}
Quinn, D. M., Chaudoir, S. R., \& Kallen, R. W. (2011). Performance and
flow: A review and integration of self-objectification research.
\emph{Self-Objectification in Women: Causes, Consequences, and
Counteractions.}, 119--138.
doi:\href{https://doi.org/10.1037/12304-006}{10.1037/12304-006}

\hypertarget{ref-quinn2011performance}{}
Quinn, D. M., Chaudoir, S. R., \& Kallen, R. W. (2011). Performance and
flow: A review and integration of self-objectification research.

\hypertarget{ref-R-base}{}
R Core Team. (2017). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-R-psych}{}
Revelle, W. (2017). \emph{Psych: Procedures for psychological,
psychometric, and personality research}. Evanston, Illinois:
Northwestern University. Retrieved from
\url{https://CRAN.R-project.org/package=psych}

\hypertarget{ref-richeson2003prejudice}{}
Richeson, J. A., \& Shelton, J. N. (2003). When prejudice does not pay:
Effects of interracial contact on executive function.
\emph{Psychological Science}, \emph{14}(3), 287--290.

\hypertarget{ref-riemer2020self}{}
Riemer, A. R., Sáez, G., Brock, R., \& Gervais, S. J. (2020).
Self-fulfilling objectification in relationships: The effects of men's
objectifying expectations on women's self-objectification during
conflict in romantic relationships. \emph{Self and Identity}, 1--7.

\hypertarget{ref-rollero2016bringing}{}
Rollero, C. (2016). Bringing objectification into social relationships
research: Is self-objectification harmful for authenticity? \emph{The
Spanish Journal of Psychology}, (19), 1--7.

\hypertarget{ref-saguyetal2010}{}
Saguy, T., Quinn, D., Dovidio, J., \& Pratto, F. (2010). Interacting
like a body: Objectification can lead women to narrow their presence in
social interactions. \emph{Psychological Science}, \emph{21}, 178--82.
doi:\href{https://doi.org/10.1177/0956797609357751}{10.1177/0956797609357751}

\hypertarget{ref-sanchez2008romance}{}
Sanchez, D. T., \& Broccoli, T. L. (2008). The romance of
self-objectification: Does priming romantic relationships induce states
of self-objectification among women? \emph{Sex Roles}, \emph{59}(7-8),
545--554.

\hypertarget{ref-shelton2005expecting}{}
Shelton, J. N., Richeson, J. A., \& Salvatore, J. (2005). Expecting to
be the target of prejudice: Implications for interethnic interactions.
\emph{Personality and Social Psychology Bulletin}, \emph{31}(9),
1189--1202.

\hypertarget{ref-spencer1999stereotype}{}
Spencer, S. J., Steele, C. M., \& Quinn, D. M. (1999). Stereotype threat
and women's math performance. \emph{Journal of Experimental Social
Psychology}, \emph{35}(1), 4--28.

\hypertarget{ref-R-apaTables}{}
Stanley, D. (2018). \emph{ApaTables: Create american psychological
association (apa) style tables}. Retrieved from
\url{https://CRAN.R-project.org/package=apaTables}

\hypertarget{ref-strelan2005women}{}
Strelan, P., \& Hargreaves, D. (2005). Women who objectify other women:
The vicious circle of objectification? \emph{Sex Roles},
\emph{52}(9-10), 707--712.

\hypertarget{ref-strelan2018birds}{}
Strelan, P., \& Pagoudis, S. (2018). Birds of a feather flock together:
The interpersonal process of objectification within intimate
heterosexual relationships. \emph{Sex Roles}, \emph{79}(1-2), 72--82.

\hypertarget{ref-taylor2018threat}{}
Taylor, V. J., Garcia, R. L., Shelton, J. N., \& Yantis, C. (2018). ``A
threat on the ground'': The consequences of witnessing
stereotype-confirming ingroup members in interracial interactions.
\emph{Cultural Diversity and Ethnic Minority Psychology}, \emph{24}(3),
319.

\hypertarget{ref-teran2020relational}{}
Terán, L., Jiao, J., \& Aubrey, J. S. (2020). The relational burden of
objectification: Exploring how past experiences of interpersonal sexual
objectification are related to relationship competencies. \emph{Sex
Roles}, 1--16.

\hypertarget{ref-tolman2006looking}{}
Tolman, D. L., Impett, E. A., Tracy, A. J., \& Michael, A. (2006).
Looking good, sounding good: Femininity ideology and adolescent girls'
mental health. \emph{Psychology of Women Quarterly}, \emph{30}(1),
85--95.

\hypertarget{ref-tylka2010integrating}{}
Tylka, T. L., \& Sabik, N. J. (2010). Integrating social comparison
theory and self-esteem within objectification theory to predict women's
disordered eating. \emph{Sex Roles}, \emph{63}(1-2), 18--31.

\hypertarget{ref-vangeel2018multidimensional}{}
Vangeel, L., Vandenbosch, L., \& Eggermont, S. (2018). The
multidimensional self-objectification process from adolescence to
emerging adulthood. \emph{Body Image}, \emph{26}, 60--69.

\hypertarget{ref-R-tidyverse}{}
Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
'tidyverse'}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyverse}

\hypertarget{ref-wollast2020cultural}{}
Wollast, R., Riemer, A. R., Gervais, S. J., Grigoryan, L., Bernard, P.,
\& Klein, O. (2020). How cultural orientation and self-compassion shape
objectified body consciousness for women from america, belgium, russia,
and thailand. \emph{Self and Identity}, 1--21.

\hypertarget{ref-R-knitr}{}
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.).
Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://yihui.name/knitr/}

\hypertarget{ref-R-rmarkdown}{}
Xie, Y., Allaire, J., \& Grolemund, G. (2018). \emph{R markdown: The
definitive guide}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved
from \url{https://bookdown.org/yihui/rmarkdown}

\hypertarget{ref-yilmaz2019whose}{}
Yilmaz, T., \& Bozo, Ö. (2019). Whose gaze is more objectifying? An
experimental study of college women's state self-objectification, body
shame, negative mood, and body dissatisfaction. \emph{Mediterranean
Journal of Clinical Psychology}, \emph{7}(2).

\hypertarget{ref-zurbriggen2011self}{}
Zurbriggen, E. L., Ramsey, L. R., \& Jaworski, B. K. (2011). Self-and
partner-objectification in romantic relationships: Associations with
media consumption and relationship satisfaction. \emph{Sex Roles},
\emph{64}(7-8), 449--462.

\endgroup

\end{document}